---
title: "ADRECMO — DPP3"
subtitle: ""
author: "KIMMOUN DUARTE"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center"
)
```

# STATISTICAL METHOD

Continuous variables are reported as median and interquartile range [IQR], whereas categorical variables are presented as counts and percentages. Baseline comparisons were performed by stratifying patients according to the median pre-implantation cDPP3 concentration into below median and above median cDPP3 groups. Considering the sample size and the exploratory, post hoc nature of this study, between-group differences are reported as Hodges–Lehmann estimates with 95% confidence intervals (derived from the Wilcoxon rank-sum test) for continuous variables, and as absolute differences in group proportions with 95% confidence intervals (estimated using the Agresti–Caffo method) for categorical variables. Continuous variables with a skewed distribution were log transformed for all subsequent models. Correlations between cDPP3 concentrations and continuous variables were evaluated using Spearman test. Variables measured at three predefined time points were compared using linear mixed-effects models with Kenward–Roger correction. Residuals of the models were checked. 30-day survival was evaluated in high and low cDPP3 groups drawing Kaplan–Meier curves. Heart transplantation and long-term mechanical circulatory support were treated as censoring events. A multivariable Cox model investigating the association between cDPP3 and the 30-day survival was performed: considering the sample size and the fact that cDPP3 is regarded as one of the main drivers highly correlated with clinico-biological severity and outcome, confounders were selected a priori and included cardiac arrest before VA-ECMO, cause of cardiogenic shock (ischemic or non-ischemic), the delay between ICU admission and VA-ECMO implantation. Model assumption was checked using Schoenfeld residuals. All statistical analyses were performed using R (version 4.5.2 (2025-10-31)). A two-sided p-value \< 0.05 was considered statistically significant and as the results were considered exploratory no adjustment for alpha multiplicity was performed.

# GLOBAL SETUP

```{r global-setup, include=FALSE}
# ---- Packages ----
library(dplyr)
library(tidyr)
library(stringr)
library(tableone)
library(flextable)
library(officer)
library(ggplot2)
library(survival)
library(survminer)
library(broom)
library(lubridate)
library(lme4)
library(lmerTest)
library(pbkrtest)
library(patchwork)
library(tibble)
library(DescTools)

# ---- Global constants ----
CAPTION_T1 <- "Table 1: Characteristics of the population by circulating DPP3 median measured at time of VA-ECMO implantation"
CAPTION_T2 <- "Table 2: Outcomes according to circulating DPP3 median measured at time of VA-ECMO implantation"
FOOTNOTE_TABLES <- "Data are expressed in n(%) for categorical variables and median [25th–75th] for continuous variables.Between-group differences are reported as Hodges–Lehmann estimates with 95% confidence intervals for continuous variables, and as absolute differences in group proportions with 95% confidence intervals (estimated using the Agresti–Caffo method) for categorical variables."

df <- readRDS("DATA/df_managed.rds")

#table_1 var
t1_vars_cont <- c(
  "Age", "BMI", "PAS_D0", "PAD_D0", "PAM_D0", "HR_D0", "Preecmo_tte_ef",
  "Preecmo_tte_vtiao", "Bili_D0", "Tropo_i_hs_D0", "Ntprobnp_D0", "Plq_D0"
)

t1_vars_cat <- c(
  "Gender", "HTN", "DM", "Immunodepression", "CKD", "Neurologic_deficit",
  "Chronic_respiratory_disease", "HF", "Alphablocker", "Betablocker",
  "cause", "ische", "Cardiac_arrest_before_canul", "Intubation"
)

t1_vars_all <- c(t1_vars_cat, t1_vars_cont)

#labels cDPP3
cutoff <- round(median(df$DPP3_D0, na.rm = TRUE))

below <- paste0("cDPP3≤", cutoff, " ng/mL")
above <- paste0("cDPP3>", cutoff, " ng/mL")

label_median <- sprintf("cDPP3 median=%.0f ng/mL",
  median(df$DPP3_D0, na.rm = TRUE)
)

#table_2 var

t2_vars_cat_bin <- c(
  "BCPIA_Tot", "EER_Tot","j30_deces",
  "Outcome_Death", "Outcome_ECMO_Weaning",
  "Outcome_Bridge_LVAD", "Outcome_Bridge_Transplant"
)
t2_vars_cont <- c("Lenght_vm", "Time_hosp", "Time_ecmo",
  "ECMO_duration_weaned", "ECMO_duration_Dead"
)
t2_vars_all <- c(t2_vars_cat_bin, t2_vars_cont)

# Data preparation
df <- df %>%
  mutate(
    #Table1
    across(all_of(t1_vars_cont), as.numeric),
    across(all_of(t1_vars_cat), as.factor),
    #table 2
    across(all_of(t2_vars_cont), as.numeric),
      j30_deces =  if_else(diff_days>=30, 0, 1),
      Outcome_Death            = Outcome == "Death",
      Outcome_ECMO_Weaning     = Outcome == "ECMO Weaning",
      Outcome_Bridge_LVAD      = Outcome == "Bridge to LVAD",
      Outcome_Bridge_Transplant= Outcome == "Bridge to Transplant",
      across(all_of(t2_vars_cat_bin), as.factor),
      #FIG 2 SURVIVAL
      diff_days_30 = if_else(diff_days >= 30, 30, diff_days),
      outcome_censored_30 = case_when(
      diff_days >= 30 & outcome_censored == 1 ~ 0,
      TRUE ~ outcome_censored),
      DPP3_group = if_else(DPP3_D0 <= cutoff,"below","above"),
      DPP3_median = factor(DPP3_group, levels = c( "above","below"), labels = c(above, below)),

      ecmo_start_date = if_else(ID == "10-HC", as.Date("2018-06-27"), ecmo_start_date),
      time_adm_ECMO = time_length(interval(icu_admiss_date, ecmo_start_date), "days"),
      Cardiac_arrest_before_canul = as.factor(Cardiac_arrest_before_canul), cause = as.factor(cause),
 #FIG 4 CORR   
      DPP3_delta_D0_D3_5 = log10(DPP3_D3_5) - log10(DPP3_D0),
      delta_Lact_D0_D3_5 = jp2_lact - Lact_D0,
      delta_NADcum_D0_D3_5 = jp2_NAd_cum_w - NAD_24H,
      delta_ASAT_D0_D3_5 = jp2_asat - ASAT_D0,
      delta_CKD_D0_D3_5 = CKD_J3J5 - CKD_D0,
 #FIG 5 HAPTO
     hapto_J0 = as.numeric(hapto_J0),
    `hapto_J3-J5` = as.numeric(`hapto_J3-J5`), delta_hapto = if_else(!is.na(hapto_J0) & !is.na(`hapto_J3-J5`),`hapto_J3-J5` - hapto_J0,NA_real_),
    hapto_trend = case_when(is.na(delta_hapto) ~ NA_character_,
      delta_hapto > 0    ~ "increase",
      TRUE               ~ "decrease"
    )
 )

#FIGURE 3  Data for longitudinal  cDPP3  course
df_dpp3_long <- df %>% 
  select(ID, DPP3_D0, DPP3_D3_5, DPP3_w, Outcome_death_bridge) %>%
  pivot_longer(cols = starts_with("DPP3_"), names_to = "timepoint_raw", values_to = "value") %>%
  mutate(
    timepoint = str_remove(timepoint_raw, "^DPP3_"),
    timepoint = case_when(
      timepoint == "D0"   ~ "implantation",
      timepoint == "D3_5" ~ "day 3 to 5",
      timepoint == "w"    ~ "explantation",
      TRUE ~ timepoint
    ),
    timepoint = factor(timepoint, levels = c("implantation", "day 3 to 5", "explantation")),
    value_log = log10(value)
  )

source("fonction_adrecmo_DPP3.R")
```

# TABLE 1: Baseline characteristics by DPP3 median

```{r table-1, include=FALSE}
df$DPP3_median
t1_tableone_obj <- CreateTableOne(
  vars       = t1_vars_all,
  factorVars = t1_vars_cat,
  strata     = "DPP3_median",
  data       = df
)

t1_tableone_print <- print(
  t1_tableone_obj,
  nonnormal      = t1_vars_cont,
  contDigits     = 1,
  showAllLevels  = FALSE,
  test          = FALSE   
)

#H-L estimates with 95%CI
hl_diff_ci <- purrr::map_dfr(
  t1_vars_cont,
  ~{
    x <- df[[.x]]
    g <- df$DPP3_median
    
    ok <- !is.na(x) & !is.na(g)
    x  <- x[ok]
    g  <- droplevels(g[ok])
    
    # si un groupe vide -> NA
    if (nlevels(g) < 2 || any(table(g) == 0)) {
      return(tibble(
        raw_id = .x,
        `Median difference [95% CI]` = NA_character_
      ))
    }
    
    wt <- suppressWarnings(
      wilcox.test(
        x ~ g,
        conf.int = TRUE,
        exact = FALSE
      )
    )
    
    # estimate = HL shift (tends to represent a "median difference" / location shift)
    est <- unname(wt$estimate)
    ci  <- unname(wt$conf.int[1:2])
    
    tibble(
      raw_id = .x,
      `Median difference [95% CI]` =
        sprintf("%.1f [%.1f; %.1f]", est, ci[1], ci[2])
    )
  }
)

# --- Différence de proportions + IC95% via BinomDiffCI (méthode scorecc ou mn) ---
diff_prop_ci <- purrr::map_dfr(
  t1_vars_cat,
  ~{
    x <- df[[.x]]
    g <- df$DPP3_median
    
    # masquage des NA
    ok <- !is.na(x) & !is.na(g)
    x  <- x[ok]
    g  <- droplevels(g[ok])
    
    # si moins de 2 niveaux ou effectifs nuls → NA
    if (nlevels(g) < 2 || any(table(g) == 0)) {
      return(tibble(
        raw_id = .x,
        `Difference in proportions (% points) [95% CI]` = NA_character_
      ))
    }
    
    # compter succès (par défaut niveau "1" ou TRUE)
    tab <- table(g, x)
    levs <- colnames(tab)
    event <- if ("1" %in% levs) "1" else levs[2]  # prendre "1" si possible, sinon second niveau
    
    x1 <- sum(x[g == "cDPP3≤73 ng/mL"] == event)
    n1 <- sum(g == "cDPP3≤73 ng/mL")
    x0 <- sum(x[g == "cDPP3>73 ng/mL"] == event)
    n0 <- sum(g == "cDPP3>73 ng/mL")
    
    # IC de différence de proportions (scorecc ou mn)
    ci <- tryCatch(
      DescTools::BinomDiffCI(
        x1 = x1, n1 = n1,
        x2 = x0, n2 = n0,
        conf.level = 0.95,
        method = "ac"     
      ),
      error = function(e) NULL
    )
    
    if (is.null(ci)) {
      return(tibble(
        raw_id = .x,
        `Difference in proportions (% points) [95% CI]` = NA_character_
      ))
    }
    
    # diff observée en pourcentage
    diff_pct  <- 100 * (x1/n1 - x0/n0)
    low_pct   <- 100 * ci[2]
    high_pct  <- 100 * ci[3]
    
    tibble(
      raw_id = .x,
      `Difference in proportions (% points) [95% CI]` =
        sprintf("%+.1f [%+.1f; %+.1f]", diff_pct, low_pct, high_pct)
    )
  }
)

df_t1_tableone <- as.data.frame(t1_tableone_print) %>%
  tibble::rownames_to_column("Variables")

df_t1_missing <- df %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(cols = everything(), names_to = "raw_id", values_to = "n_missing")


df_t1_tableone_id <- df_t1_tableone %>%
  mutate(
    raw_id = sub(
      "\\(median \\[IQR\\]\\)|= 1 \\(\\%\\)|= Women \\(\\%\\)|\\(\\%\\)|= Non ischémique \\(\\%\\)|= ischémique \\(\\%\\)",
      "",
      Variables
    ),
    raw_id = trimws(raw_id),
    raw_id = sub(" = YES", "", raw_id),
    Variables = str_remove(Variables, " \\(median \\[IQR\\]\\)")
  )


df_t1_final <- df_t1_tableone_id %>%
  left_join(df_t1_missing, by = "raw_id") %>%
  left_join(hl_diff_ci,   by = "raw_id") %>%
  left_join(diff_prop_ci, by = "raw_id") %>%
  mutate(
    `Group difference [95% CI]` = dplyr::if_else(
      !is.na(`Median difference [95% CI]`),
      `Median difference [95% CI]`,
      `Difference in proportions (% points) [95% CI]`
    )
  ) %>%
  transmute(
    Variables,
    `cDPP3≤73 ng/mL`,
    `cDPP3>73 ng/mL`,
    Missing = n_missing,
    `Group difference [95% CI]`
  )

#ARRONDIR#
zerodigit <- df_t1_tableone %>%
  dplyr::filter(!grepl("^Lact_D0 ", Variables)) %>%     
  dplyr::pull(Variables)

rows_to_round <- df_t1_tableone$Variables %in% zerodigit

df_t1_tableone$`cDPP3≤73 ng/mL`[rows_to_round] <-
  round_nonnormal_0(df_t1_tableone$`cDPP3≤73 ng/mL`[rows_to_round])

df_t1_tableone$`cDPP3>73 ng/mL`[rows_to_round] <-
  round_nonnormal_0(df_t1_tableone$`cDPP3>73 ng/mL`[rows_to_round])


df_t1_final$`Group difference [95% CI]` <-
  round_nonnormal_0(df_t1_final$`Group difference [95% CI]`)

dict_t1_labels <- c(
  "n" = "n",
  "Gender = Women (%)" = "Gender = Female (%)",
  "HTN = 1 (%)" = "Hypertension (%)",
  "DM = 1 (%)" = "Diabetes Mellitus (%)",
  "Immunodepression = 1 (%)" = "Immunosuppression (%)",
  "CKD = 1 (%)" = "Chronic Kidney Disease (%)",
  "Neurologic_deficit = 1 (%)" = "Neurological Deficit (%)",
  "Chronic_respiratory_disease = 1 (%)" = "Chronic Respiratory Disease (%)",
  "HF = 1 (%)" = "Heart Failure (%)",
  "Alphablocker = 1 (%)" = "Alpha-blocker (%)",
  "Betablocker = YES (%)" = "Beta-blocker (%)",
  "cause = Non ischémique (%)" = "Non-ischemic (%)",
  "ische = ischémique (%)" = "Ischemic (%)",
  "Cardiac_arrest_before_canul = 1 (%)" = "Cardiac Arrest Before Cannulation (%)",
  "Intubation = 1 (%)" = "Mechanical ventilation (%)",
  "Age" = "Age (years)",
  "BMI" = "Body Mass Index (kg/m²)",
  "PAS_D0" = "Systolic Blood Pressure (mmHg)",
  "PAD_D0" = "Diastolic Blood Pressure (mmHg)",
  "PAM_D0" = "Mean Arterial Pressure (mmHg)",
  "HR_D0" = "Heart Rate (bpm)",
  "Preecmo_tte_ef" = "Ejection Fraction (%)",
  "Preecmo_tte_vtiao" = "Aortic Velocity Time Integral (cm)",
  "Bili_D0" = "Bilirubin (mg/dL)",
  "Tropo_i_hs_D0" = "High Sensitivity Troponin I (ng/L)",
  "Ntprobnp_D0" = "NT-proBNP (pg/mL)",
  "Plq_D0" = "Platelet Count (10⁹/L)"
)

df_t1_final <- df_t1_final %>%
  mutate(Variables = dplyr::recode(Variables, !!!dict_t1_labels))

t1_sections <- tibble::tribble(
  ~Variables, ~Missing, ~`cDPP3>73 ng/mL`, ~`cDPP3≤73 ng/mL`, ~`Group difference [95% CI]`,
  "Demographics", NA_integer_, "", "", "",
  "Medical History", NA_integer_, "", "", "",
  "Prior-adrenergic medications", NA_integer_, "", "", "",
  "Cardiogenic shock cause", NA_integer_, "", "", "",
  "Clinical parameters", NA_integer_, "", "", "",
  "Heart function", NA_integer_, "", "", "",
  "Supports", NA_integer_, "", "", "",
  "Biological assessment", NA_integer_, "", "", ""
)

t1_order <- c(
  "n",
  "Demographics",
  "Age (years)",
  "Gender = Female (%)",
  "Body Mass Index (kg/m²)",
  "Medical History",
  "Hypertension (%)",
  "Diabetes Mellitus (%)",
  "Heart Failure (%)",
  "Chronic Kidney Disease (%)",
  "Chronic Respiratory Disease (%)",
  "Neurological Deficit (%)",
  "Immunosuppression (%)",
  "Prior-adrenergic medications",
  "Beta-blocker (%)",
  "Alpha-blocker (%)",
  "Cardiogenic shock cause",
  "Non-ischemic (%)",
  "Ischemic (%)",
  "Cardiac Arrest Before Cannulation (%)",
  "Clinical parameters",
  "Systolic Blood Pressure (mmHg)",
  "Diastolic Blood Pressure (mmHg)",
  "Mean Arterial Pressure (mmHg)",
  "Heart Rate (bpm)",
  "heart function",
  "Ejection Fraction (%)",
  "Aortic Velocity Time Integral (cm)",
  "supports",
  "Mechanical ventilation (%)",
  "Biological assessment",
  "Bilirubin (mg/dL)",
  "High Sensitivity Troponin I (ng/L)",
  "NT-proBNP (pg/mL)",
  "Platelet Count (10⁹/L)"
)

df_t1_display <- bind_rows(t1_sections, df_t1_final) %>%
  mutate(Variables = factor(Variables, levels = t1_order)) %>%
  arrange(Variables) %>%
  mutate(    across(everything(), ~gsub("\\s*[;,]\\s*", " – ", .x)),
    is_section = is.na(Missing) & `cDPP3≤73 ng/mL` == "" & `cDPP3>73 ng/mL` == "")

ft_t1 <- flextable(df_t1_display %>% select(-is_section)) %>%
  set_caption(CAPTION_T1) %>%
  autofit() %>%
  align(j = 2:5, align = "center", part = "body") %>%
  bold(i = which(df_t1_display$is_section), bold = TRUE) %>%
  align(i = which(df_t1_display$is_section), j = 1, align = "left", part = "body") %>%
  bg(i = which(df_t1_display$is_section), bg = "#FFFFFF", part = "body") %>%
  hline_bottom(part = "header", border = fp_border(color = "black", width = 1)) %>%
  hline_bottom(part = "body", border = fp_border(color = "black", width = 1)) %>%
  add_footer_lines(values = FOOTNOTE_TABLES) %>%
  align(part = "footer", align = "left") %>%
  fontsize(part = "footer", size = 9)

save_as_docx(ft_t1, path = "Tables/Table1.docx")
```

```{r table 1 print}

ft_t1
```

# TABLE 2: Outcome according to cDDP3 median

```{r table-2, include=FALSE}

t2_tableone_obj <- CreateTableOne(
  vars       = t2_vars_all,
  factorVars = t2_vars_cat_bin,
  strata     = "DPP3_median",
  data       = df
)

t2_tableone_print <- print(
  t2_tableone_obj,
  nonnormal      = t2_vars_cont,
  contDigits     = 1,
  showAllLevels  = FALSE,
  test = FALSE
  
)

 #Hodges–Lehmann for continuous outcomes
hl_diff_ci_t2 <- purrr::map_dfr(
  t2_vars_cont,
  ~{
    x <- df[[.x]]
    g <- df$DPP3_median
    
    ok <- !is.na(x) & !is.na(g)
    x  <- x[ok]
    g  <- droplevels(g[ok])
    
    if (nlevels(g) < 2 || any(table(g) == 0)) {
      return(tibble(raw_id = .x, Group_diff_CI = NA_character_))
    }
    
    wt <- suppressWarnings(
      wilcox.test(x ~ g, conf.int = TRUE, exact = FALSE)
    )
    est <- unname(wt$estimate)
    ci  <- unname(wt$conf.int[1:2])
    
    tibble(
      raw_id = .x,
      Group_diff_CI = sprintf("%.1f [%.1f; %.1f]", est, ci[1], ci[2])
    )
  }
)

# Difference in proportions for categorical outcomes
prop_diff_ci_t2 <- purrr::map_dfr(
  t2_vars_cat_bin,
  ~{
    x <- df[[.x]]
    g <- df$DPP3_median
    
    ok <- !is.na(x) & !is.na(g)
    x  <- x[ok]
    g  <- droplevels(g[ok])
    
    if (nlevels(g) < 2 || any(table(g) == 0)) {
      return(tibble(raw_id = .x, Group_diff_CI = NA_character_))
    }
    
    x <- as.factor(x)
    levs <- levels(x)
    event <- if ("1" %in% levs) "1" else levs[2]
    
    x1 <- sum(x[g == "cDPP3>73 ng/mL"] == event)
    n1 <- sum(g == "cDPP3>73 ng/mL")
    x0 <- sum(x[g == "cDPP3≤73 ng/mL"] == event)
    n0 <- sum(g == "cDPP3≤73 ng/mL")
    
    ci <- tryCatch(
      DescTools::BinomDiffCI(
        x1 = x1, n1 = n1,
        x2 = x0, n2 = n0,
        conf.level = 0.95,
        method = "ac"
      ),
      error = function(e) NULL
    )
    
    if (is.null(ci)) {
      Group_diff_CI <- NA_character_
    } else {
      diff_pct <- 100 * (x1/n1 - x0/n0)
      low_pct  <- 100 * ci[2]
      high_pct <- 100 * ci[3]
      Group_diff_CI <- sprintf("%+.1f [%+.1f; %+.1f]",
                               diff_pct, low_pct, high_pct)
    }
    
    tibble(raw_id = .x, Group_diff_CI = Group_diff_CI)
  }
)


df_t2_final <- as.data.frame(t2_tableone_print) %>%
  tibble::rownames_to_column("Variables") %>%
  mutate(Variables = str_remove(Variables, " \\(median \\[IQR\\]\\)")) %>%
  mutate(raw_id = trimws(sub("\\(.*\\)|=.*", "", Variables))) %>%
  left_join(hl_diff_ci_t2,   by = "raw_id") %>%
  left_join(prop_diff_ci_t2, by = "raw_id") %>%
  mutate(
    `Group difference [95% CI]` = dplyr::coalesce(Group_diff_CI.x, Group_diff_CI.y)
  ) %>%
  transmute(
    Variables,
    `cDPP3≤73 ng/mL`,
    `cDPP3>73 ng/mL`,
    `Group difference [95% CI]`
  )



df_t2_final$`cDPP3≤73 ng/mL` <-
  round_nonnormal_0(df_t2_final$`cDPP3≤73 ng/mL`)

df_t2_final$`cDPP3>73 ng/mL`<-
  round_nonnormal_0(df_t2_final$`cDPP3>73 ng/mL`)

df_t2_final$`Group difference [95% CI]` <-
  round_nonnormal_0(df_t2_final$`Group difference [95% CI]`)


t2_sections <- tibble::tribble(
  ~Variables,  ~`cDPP3>73 ng/mL`,~`cDPP3≤73 ng/mL`, ~`Group difference [95% CI]`,
  "Outcome", "", "", ""
)

dict_t2_labels <- c(
  "n"                                     = "n",
  "Outcome_Bridge_LVAD = TRUE (%)"        = "Bridge to LVAD (%)",
  "Outcome_Bridge_Transplant = TRUE (%)"  = "Bridge to Transplant (%)",
  "Outcome_Death = TRUE (%)"                 = "Death in ICU (%)",
  "Outcome_ECMO_Weaning = TRUE (%)"          = "ECMO Weaning (%)",
  "BCPIA_Tot = 1 (%)"        = "Intra-Aortic Balloon Pump (%)",
  "j30_deces = 1 (%)"        = "30-day Mortality (%)",
  "EER_Tot = 1 (%)"          = "Renal Replacement Therapy (%)",
  "Lenght_vm"                = "Duration of Mechanical Ventilation (days)",
  "Time_hosp"                = "ICU Length of Stay (days)",
  "Time_ecmo"                = "ECMO Duration (days)",
  "ECMO_duration_weaned"     = "ECMO Duration in weaned (days)",
  "ECMO_duration_Dead"       = "ECMO Duration in deceased (days)"
)

t2_order <- c(
  "n",
  "ECMO Duration (days)",
  "ECMO Duration in weaned (days)",
  "ECMO Duration in deceased (days)",
  "Intra-Aortic Balloon Pump (%)",
  "Renal Replacement Therapy (%)",
  "Duration of Mechanical Ventilation (days)",
  "ICU Length of Stay (days)",
  "Outcome",
  "Death in ICU (%)",
  "Bridge to LVAD (%)",
  "Bridge to Transplant (%)",
  "ECMO Weaning (%)",
  "30-day Mortality (%)"
)


df_t2_display <- bind_rows(t2_sections, df_t2_final) %>%
  mutate(   across(everything(), ~gsub("\\s*[;,]\\s*", " – ", .x)),
    Variables = dplyr::recode(Variables, !!!dict_t2_labels),
    Variables = factor(
      Variables,
      levels = c("n", setdiff(t2_order, "n"))
    )
  ) %>%
  arrange(Variables)


ft_t2 <- flextable(df_t2_display) %>%
  set_caption(CAPTION_T2) %>%
  autofit() %>%
  align(j = 2:4, align = "center", part = "body") %>%
  bold(i = ~ Variables == "Outcome", bold = TRUE) %>%
  align(i = ~ Variables == "Outcome", j = 1, align = "left", part = "body") %>%
  hline_bottom(part = "header", border = fp_border(color = "black", width = 1)) %>%
  hline_bottom(part = "body", border = fp_border(color = "black", width = 1)) %>%
  add_footer_lines(values = FOOTNOTE_TABLES) %>%
  align(part = "footer", align = "left") %>%
  fontsize(part = "footer", size = 9)

save_as_docx(ft_t2, path = "Tables/Table2.docx")
```

```{r table 2 print}

ft_t2
```

# FIGURE 1: Association between circulating DPP3 concentrations and organ dysfunction at VA-ECMO implantation. Patients were stratified according to baseline circulating DPP3 (cDPP3) concentrations using the median value (73 ng/mL).

```{r fig-1, include=FALSE}

####Values of the variables

fig1_vars_cont <- c("sofa_D0","Lact_D0","NAD_24H","CKD_D0","j0_ecmo_debit","ASAT_D0")

fig1_vars_all <- fig1_vars_cont

fig1_tableone_obj <- CreateTableOne(
  vars       = fig1_vars_all,
  factorVars = NULL,
  strata     = "DPP3_median",
  data       = df
)

fig1_tableone_print <- print(
  fig1_tableone_obj,
  nonnormal      = fig1_vars_cont,
  contDigits     = 1,
  showAllLevels  = FALSE,
  test          = T
)

# Create plots
p_dpp3_sofa <- plot_box_by_dpp3_group(df, "sofa_D0", "SOFA")
p_dpp3_lact <- plot_box_by_dpp3_group(df, "Lact_D0", "Lactate\n(mmol/L)")
p_dpp3_nad  <- plot_box_by_dpp3_group(df, "NADcum_D0_w", "Cumulated norepinephrine dose\nday of ECMO implantation (γ/kg)")
p_dpp3_ckd  <- plot_box_by_dpp3_group(df, "CKD_D0", "eGFR\n(mL/min/1.73m²)")
p_dpp3_flow <- plot_box_by_dpp3_group(df, "j0_ecmo_debit", "ECMO Flow\n(L/min)", x_lab = "cDPP3")
p_dpp3_asat <- plot_box_by_dpp3_group(df, "ASAT_D0", "AST\n(IU/L)", x_lab = "cDPP3")

# Combine plots into a single figure
fig_dpp3_baseline <- (p_dpp3_sofa + p_dpp3_lact) /
  (p_dpp3_nad + p_dpp3_flow) /
  (p_dpp3_ckd+ p_dpp3_asat) +
  plot_annotation(
    title = sprintf("cDPP3 median at time of VA-ECMO implantation=%.0fng/mL",
      median(df$DPP3_D0, na.rm = TRUE)
    ),
    tag_levels = c("A", "B", "C", "D", "E", "F")
  ) & theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))



# Save the figure
ggsave(
  "FIGURES/Figure 1.pdf",
  plot   = fig_dpp3_baseline,
  width  = 160, height = 200, units = "mm",
  device = cairo_pdf, bg = "white", scale = 1.5
)


```

```{r Fig-1 print, echo=F, out.width="80%", fig.align="center", fig.cap="FIGURE 1: Association between circulating DPP3 concentrations and organ dysfunction at VA-ECMO implantation. Patients were stratified according to baseline circulating DPP3 (cDPP3) concentrations using the median value (73 ng/mL).",fig.width=8, fig.height=10}

fig_dpp3_baseline
```

# FIGURE 2: Association between baseline circulating DPP3 concentrations measured at VA-ECMO implantation and 30-day survival in patients with cardiogenic shock supported with VA-ECMO.

```{r fig-2, include=FALSE}

km_fit_30d <- survfit(Surv(diff_days_30, outcome_censored_30) ~ DPP3_median, data = df)

cox_uni <- coxph(Surv(diff_days_30, outcome_censored_30) ~ relevel(DPP3_median, ref = "cDPP3≤73 ng/mL"), data = df)

cox_mv <- coxph(Surv(diff_days_30, outcome_censored_30) ~ relevel(DPP3_median, ref = "cDPP3≤73 ng/mL") + time_adm_ECMO + cause + Cardiac_arrest_before_canul, data = df)

cox_uni_tidy <- tidy(cox_uni, exponentiate = TRUE, conf.int = TRUE)
cox_mv_tidy <- tidy(cox_mv, exponentiate = TRUE, conf.int = TRUE)

lab_uni <- cox_uni_tidy %>%
  mutate(label = sprintf("Unadjusted HR=%.2f (95%% CI %.2f–%.2f)", estimate, conf.low, conf.high)) %>%
  pull(label)

lab_adj <- cox_mv_tidy %>%
  filter(term == "relevel(DPP3_median, ref = \"cDPP3≤73 ng/mL\")cDPP3>73 ng/mL") %>%
  mutate(label = sprintf("Adjusted HR=%.2f (95%% CI %.2f–%.2f)", estimate, conf.low, conf.high)) %>%
  pull(label)

lab_surv <- paste(lab_uni, lab_adj, sep = "\n")

ph_test_mv <- cox.zph(cox_mv)

surv_plot <- ggsurvplot(
  km_fit_30d,
  data = df,
  pval = FALSE,
  conf.int = FALSE,
  risk.table = TRUE,
  risk.table.height = 0.2,
  risk.table.y.text = TRUE,
  risk.table.title = "Number at Risk",
  xlim = c(0, 30),
  ylab = "30-day survival",
  break.time.by = 5,
  ggtheme = theme_classic(),
  surv.scale = "percent",
  legend.labs = c(below, above),
  legend = "none"
)

surv_plot$plot <- surv_plot$plot +
  annotate("text",
           x = 5,
           y = 0.15,
           label = lab_surv,
           hjust = 0, size = 4) +
  annotate("text",
           x = 15,
    y = 1,
    label = label_median,
    hjust = 0.5,
    size = 4,
    fontface = "bold"
  )

fig_survival <- ggpubr::ggarrange(
  surv_plot$plot,
  surv_plot$table,
  ncol = 1, nrow = 2,
  heights = c(3, 1)
)

ggsave(
  "FIGURES/Figure 2.pdf",
  plot   = fig_survival,
  width  = 180, height = 140, units = "mm",
  device = cairo_pdf, bg = "white"
)
```

```{r Fig-2, echo=F, out.width="80%", fig.align="center", fig.cap="FIGURE 2: Association between baseline circulating DPP3 concentrations measured at VA-ECMo implantation and 30-day survival in patients with cardiogenic shock supported with VA-ECMO.",fig.width=8, fig.height=10}

fig_survival
```

# FIGURE 3: Time course of circulating DPP3 concentrations in patients with cardiogenic shock supported with VA-ECMO.

```{r Fig-3, include=FALSE}

df_dpp3_n_time <- df_dpp3_long %>%
  group_by(timepoint) %>%
  summarise(n = sum(!is.na(value)), 
            median = median(value, na.rm = TRUE),
            IQR_low = quantile(value, 0.25, na.rm = TRUE),
            IQR_high = quantile(value, 0.75, na.rm = TRUE),.groups = "drop")

#stat
lmm_dpp3_time <- lmer(value_log ~ timepoint + (1 | ID), data = df_dpp3_long)
anova_dpp3_time <- anova(lmm_dpp3_time, ddf = "Kenward-Roger")
p_time_dpp3 <- format.pval(anova_dpp3_time["timepoint", "Pr(>F)"], digits = 3, eps = 0.001)

#graphs
p_dpp3_time <- ggplot(df_dpp3_long, aes(x = timepoint, y = value_log)) +
  geom_boxplot(outlier.shape = NA) +
  geom_text(
    data = df_dpp3_n_time,
    aes(x = timepoint, y = min(df_dpp3_long$value_log, na.rm = TRUE), label = paste0("n=", n)),
    inherit.aes = FALSE,
    vjust = 1, size = 4
  ) +
  annotate(
    "text",
    x = 1.5,
    y = max(df_dpp3_long$value_log, na.rm = TRUE) * 1.05,
    label = paste0("Time p=", p_time_dpp3),
    hjust = 0, size = 4
  ) +
  labs(x = "", y = "DPP3 (log scale)") +
  theme_minimal()+
  theme(legend.position="none",
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 14, face = "bold"))

df_dpp3_n_by_outcome <- df_dpp3_long %>%
  group_by(timepoint, Outcome_death_bridge) %>%
  summarise(n = sum(!is.na(value)), .groups = "drop")


#stat 
lmm_dpp3_inter <- lmer(value_log ~ timepoint * Outcome_death_bridge + (1 | ID), data = df_dpp3_long)
anova_dpp3_inter <- anova(lmm_dpp3_inter, ddf = "Kenward-Roger")

#hypothesis
res <- resid(lmm_dpp3_inter)
fitted <- fitted(lmm_dpp3_inter)

df_res <- data.frame(
  resid = res,
  fitted = fitted
)

ggplot(df_res, aes(x = fitted, y = resid)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(x = "Fitted values", y = "Residuals")

qqnorm(res)
qqline(res, col = "red", lwd = 2)

ggplot(df_res, aes(x = resid)) +
  geom_histogram(bins = 30, fill = "grey80", color = "black") +
  theme_minimal()
df_res <- model.frame(lmm_dpp3_inter)
df_res$resid <- resid(lmm_dpp3_inter)
df_res$fitted <- fitted(lmm_dpp3_inter)

ggplot(df_res, aes(x = timepoint, y = resid)) +
  geom_boxplot() +
  theme_minimal() +
  labs(y = "Residuals")

library(performance)

performance::check_model(lmm_dpp3_inter)
#model pas si mauvais



df_dpp3_pvals <- anova_dpp3_inter %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  transmute(
    term,
    p_value = `Pr(>F)`,
    p_label = case_when(
      p_value < 0.001 ~ "<0.001",
      TRUE ~ formatC(p_value, format = "f", digits = 2)
    ),
    label = case_when(
      term == "timepoint" ~ paste0("Time: p", p_label),
      term == "Outcome_death_bridge" ~ paste0("Outcome: p=", p_label),
      term == "timepoint:Outcome_death_bridge" ~ paste0("Interaction: p=", p_label),
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(label))

#graph by outcome
p_dpp3_by_outcome <- ggplot(df_dpp3_long, aes(x = timepoint, y = value_log, fill = factor(Outcome_death_bridge))) +
  geom_boxplot() +
  geom_text(
    data = df_dpp3_n_by_outcome,
    aes(x = timepoint, y = min(df_dpp3_long$value_log, na.rm = TRUE), label = paste0("n=", n), group = factor(Outcome_death_bridge)),
    position = position_dodge(0.7),
    inherit.aes = FALSE,
    vjust = 1, size = 4
  ) +
  annotate(
    "label",
    x = Inf, y = Inf,
    label =  paste(df_dpp3_pvals$label, collapse ="\n"),
    hjust = 1.05, vjust = 1.1)+
  labs(x = "Timepoints", y = "DPP3 (log scale)", color = "Outcomes") +
  theme_minimal()+
  theme(legend.position="bottom",
        legend.title = element_blank(),
        axis.title.x = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.x = element_text(size = 14, face = "bold")
  )
  

#combine
fig_dpp3_serial <- p_dpp3_time / p_dpp3_by_outcome +
  plot_annotation(tag_levels = c("A", "B"))

ggsave(
  "FIGURES/Figure 3.pdf",
  plot   = fig_dpp3_serial,
  width  = 160, height = 220, units = "mm",
  device = cairo_pdf, bg = "white", scale = 1.1
)
```

```{r Fig-3 print, echo=F, out.width="80%", fig.align="center", fig.cap="FIGURE 3: Time course of circulating DPP3 concentrations in patients with cardiogenic shock supported with VA-ECMO.",width=800, height=900}
fig_dpp3_serial
```

# FIGURE 4: Associations between changes in circulating DPP3 concentrations and changes in markers of organ dysfunction from VA-ECMO implantation to day 3 to 5 while on VA-ECMO.

```{r Fig-4, include=F}
  p_corr_NAD <- plot_spearman(
    df,
    "DPP3_delta_D0_D3_5",
    "delta_NADcum_D0_D3_5",
    "",
    "Δ Cumulated norepinephrine dose\n(γ/kg)"
  )
  
  p_corr_Lact <- plot_spearman(
    df,
    "DPP3_delta_D0_D3_5",
    "delta_Lact_D0_D3_5",
    "",
    "Δ Lactate\n(mmol/L)"
  )
  
  
  p_corr_ASAT <- plot_spearman(
    df,
    "DPP3_delta_D0_D3_5",
    "delta_ASAT_D0_D3_5",
    "Δ cDPP3\n(log scale)",
    "Δ ASAT\n(IU/L)"
  )
  
  p_corr_CKD <- plot_spearman(
    df,
    "DPP3_delta_D0_D3_5",
    "delta_CKD_D0_D3_5",
    "Δ cDPP3\n(log scale)",
    "Δ eGFR\n(mL/min/1.73m²)"
  )
  
  Figure_correlation_DPP3 <- (p_corr_NAD + p_corr_Lact) /
    (p_corr_ASAT + p_corr_CKD) +
    plot_annotation(title = "",
                    tag_levels = c("A","B","C","D")) & theme(plot.title=element_text(hjust=0.5))
  
  ggsave("FIGURES/Figure 4.pdf",
         plot   = Figure_correlation_DPP3,
         width  = 160, height = 200, units = "mm",
         device = cairo_pdf, bg = "white",scale=1)

```

```{r Fig-4 print, echo=F, out.width="80%", fig.align="center", fig.cap="FIGURE 4 : Associations between changes in circulating DPP3 concentrations and changes in markers of organ dysfunction from VA-ECMO implantation to day 3 to 5 while on VA-ECMO.",fig.width=8, fig.height=10}

Figure_correlation_DPP3
```

# FIGURE 5: Individual trajectories of circulating DPP3 concentrations according to haptoglobin dynamics from VA-ECMO implantation to day 3 to 5 while on VA-ECMO.

```{r Fig-5, include=F}

df_cDPP3_plot <- df_dpp3_long %>%
  filter(timepoint %in% c("implantation", "day 3 to 5")) %>%
  mutate(timepoint = factor(timepoint, levels = c("implantation", "day 3 to 5"))) %>%
  arrange(ID, timepoint) %>%
  group_by(ID) %>%
  mutate(
    delta_cDPP3 = value[timepoint == "day 3 to 5"] -
      value[timepoint == "implantation"],
    cdpp3_trend = if_else(delta_cDPP3>0, "increase", "decrease")
  ) %>%
  ungroup() %>%
  left_join(df, by = "ID")%>%
  filter(!is.na(cdpp3_trend), !is.na(hapto_trend))
df_cDPP3_plot$`hapto_J3-J5`
df_cDPP3_plot$`hapto_J0`

df_cDPP3_plot%>%
  group_by(cdpp3_trend,hapto_trend) %>%
  summarise(n = n_distinct(ID), .groups = "drop")

###test statistiques d'interaction####
mod <- lmer(value_log ~ timepoint * hapto_trend + (1 | ID), data = df_cDPP3_plot)
anova_hapto_mod <- anova(mod, ddf = "Kenward-Roger")

p_inter <- format.pval(
  anova_hapto_mod["timepoint:hapto_trend", "Pr(>F)"],
  digits = 3, eps = 0.001)

df_panels <- bind_rows(
  
  # Panel 1 — Global (all patients)
  df_cDPP3_plot %>%
    dplyr::mutate(panel = "Global"),
  
  # Panel 2 — Hapto decrease
  df_cDPP3_plot %>%
    filter(hapto_trend == "decrease") %>%
    mutate(panel = "Haptoglobin decrease"),
  
  # Panel 3 — Hapto increase
  df_cDPP3_plot %>%
    filter(hapto_trend == "increase") %>%
    mutate(panel = "Haptoglobin increase"),
  
  # Panel 4 — Hapto decrease + cDPP3 increase
  df_cDPP3_plot %>%
    filter(hapto_trend == "decrease", cdpp3_trend == "increase") %>%
    mutate(panel = "Haptoglobin decrease\n+ cDPP3 increase")
)


annot_df <- data.frame(
  panel = "Global",
  x = 1.5,
  y = max(df_panels$value_log, na.rm = TRUE) * 1.05,
  label = paste0("Interaction p=", p_inter)
)

hapto_stats <- df_cDPP3_plot %>%
  dplyr::select(-hapto_JS, -`hapto_trend`) %>%
  pivot_longer(
    cols = starts_with("hapto_"),
    names_to = "timepoint_hapto",
    values_to = "hapto_value"
  ) %>%
  dplyr::group_by(timepoint_hapto) %>%
  dplyr::summarise(
    med = median(hapto_value, na.rm = TRUE),
    q1  = quantile(hapto_value, 0.25, na.rm = TRUE),
    q3  = quantile(hapto_value, 0.75, na.rm = TRUE)
  ) %>%
  dplyr::mutate(
    label = paste0(
      round(med, 2), " (",
      round(q1, 2), " – ",
      round(q3, 2), ")"
    )
  )

hapto_map <- hapto_stats %>%
  mutate(
    timepoint_hapto = case_when(
      timepoint_hapto == "hapto_J0"     ~ "implantation",
      timepoint_hapto == "hapto_J3-J5"  ~ "day 3 to 5"
    ))%>%
  select(timepoint_hapto, label)


df_panels <- df_panels %>%
  mutate(
    timepoint_x = factor(case_when(
      panel == "Global" & timepoint == "implantation" ~ paste0("implantation","\n", 
       "haptoglobin:\n", hapto_map$label[hapto_map$timepoint_hapto == "implantation"], "\nng/mL"),
      
      panel == "Global" & timepoint == "day 3 to 5" ~paste0("day 3 to 5","\n", 
                                                            "haptoglobin:\n",  hapto_map$label[hapto_map$timepoint_hapto == "day 3 to 5"], "\nng/mL"),
      
      TRUE ~ timepoint
    )))%>%
  mutate(
    timepoint_x = factor(
      timepoint_x,
      levels = c(
        "implantation\nhaptoglobin:\n1.01 (0.16 – 1.73)\nng/mL",
        "day 3 to 5\nhaptoglobin:\n0.66 (0 – 1.51)\nng/mL",
        "implantation",
        "day 3 to 5"
      )
    )
  ) %>%
  droplevels()

panel_letters <- data.frame(
  panel = levels(factor(df_panels$panel)),
  label = LETTERS[seq_along(levels(factor(df_panels$panel)))],
  x = -Inf,
  y = Inf
)

spagh_plot <- ggplot(
  df_panels,
  aes(
    x = timepoint_x,
    y = value_log,
    group = ID,
    color = hapto_trend
  )
) +
  geom_line() +
  geom_point(size = 2) +
  
  scale_color_manual(
    values = c(
      "increase" = "#E63946",
      "decrease" = "#457B9D"
    ),
    labels = c(
      "increase" = "haptoglobin increase",
      "decrease" = "haptoglobin decrease"
    )
  ) +
  scale_y_continuous(limits = c(min(df_panels$value_log), max(df_panels$value_log)+0.2)
  ) +
  
  facet_wrap(~ panel, nrow = 1, scales = "free")+  
  labs(
    x = "",
    y = "cDPP3 (log scale)",
    color = ""
  ) +
  
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    legend.position = "bottom",
    strip.text = element_text(size = 13, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.x  = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )+
  geom_text(
    data = annot_df,
    aes(x = x, y = 3.1, label = label),
    inherit.aes = FALSE,
    hjust = 0.5,
    size = 4,
    fontface = "bold"
  )+
  geom_text(
    data = panel_letters,
    aes(x = x, y = y, label = label),
    inherit.aes = FALSE,
    hjust = -0.2,
    vjust = 1.2,
    size = 5,
    fontface = "bold"
  )


ggsave(
  "FIGURES/Figure 5.pdf",
  plot   = spagh_plot,
  width  = 200, height = 120, units = "mm",
  device = cairo_pdf, bg = "white", scale = 2)  
```

```{r Fig-5 print, echo=F, fig.align="center", fig.cap="FIGURE 5 : Individual trajectories of circulating DPP3 concentrations according to haptoglobin dynamics from VA-ECMO implantation to day 3 to 5 while on VA-ECMO. ",fig.width=14, fig.height=5}
spagh_plot
```
