---
title: "ADRECMO"
author: "Jolie Bruno, Antoine Kimmoun, Kevin Duarte, Feriel Azibani, Benjamin Deniaun"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    latex_engine: pdflatex
  word_document: null
editor_options:
  chunk_output_type: console
always_allow_html: yes
---

```{=html}
<style type="text/css">

h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  text-align: center;
}
</style>
```


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
# markdown options depending on the output (html with R code and Word without)
output <- knitr::opts_knit$get("rmarkdown.pandoc.to")
if (output=="html") knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo = TRUE, comment = "", fig.align = "center", fig.width= 17, fig.asp =0.70, out.width = "90%")
if (output=="docx") knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo = FALSE, comment = "", fig.align = "center", fig.width= 17, fig.asp =0.70, out.width = "90%")
if (output=="pdf") knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo = FALSE, comment = "", fig.align = "center", fig.width= 17, fig.asp =0.70, out.width = "90%")
options(max.print = .Machine$integer.max)
```

# Packages needed

```{r,include=FALSE}
library(dplyr)
library(lubridate)
library(gtsummary)
library(readr)
library(ggplot2)
library(tableone)
library(flextable)
library(sjlabelled)
library(readxl)
library(stringr)
library(tidyr)
library(purrr)
library(tableone)
library(dplyr)
library(naniar)
library(tidyr)
library(survival)
library(survminer)
library(patchwork)
library(ggpmisc)
library(cowplot)  # Per combinare i grafici
```

# Fonctions

```{r,include=FALSE}
source("FUNCTIONS/Examples - RCS effect + interaction cont x factor + check normality 181224.R")
```

# DATA Load

```{r,include=FALSE}
fichier <- "DATA/df_i_stabilized.xlsx"
feuilles <- excel_sheets(fichier)
data_list <- lapply(feuilles, function(feuille) {
  read_excel(fichier, sheet = feuille)
})


ADR1 <- data_list[[1]] %>%
  select( "Patient","Jour", "%Total")%>%
  filter(!is.na(`%Total`))%>%
  rename(value = `%Total`)%>%
  mutate(ID =str_replace(Patient, "Patient (\\d+) (\\w+)", "\\1-\\2"))%>%
  pivot_wider(
    names_from = "Jour",
     names_prefix = "ADR1_",
    values_from = "value",
    values_fn = mean,  # Moyenne des doublons
    values_fill = NA
  )%>%
  mutate (`ADR1_J3-5` = case_when(is.na(`ADR1_J3-5`)&!is.na(ADR1_J5)~ADR1_J5,
                             T~`ADR1_J3-5`),
          ADR1_JS = case_when(is.na(ADR1_JS)&!is.na(`ADR1_J3-5-S`)~`ADR1_J3-5-S`,
                             T~ADR1_JS))%>%
  rename(ADR1_J3_J5=`ADR1_J3-5`)%>%
  
  dplyr::select(ID,
                ADR1_J0,
                ADR1_J3_J5,
                ADR1_JS)


ADR2 <- data_list[[2]] %>%
  select( "Patient","Jour", "%Total")%>%
  filter(!is.na(`%Total`))%>%
  rename(value = `%Total`)%>%
  mutate(ID = str_replace(Patient, "Patient (\\d+) (\\w+)", "\\1-\\2"))%>%
  pivot_wider(
    names_from = "Jour",
    values_from = "value",
     names_prefix = "ADR2_" ,
    values_fn = mean,  # Moyenne des doublons
    values_fill = NA
  )%>%
  mutate (
          ADR2_JS = case_when(is.na(ADR2_JS)&!is.na(`ADR2_J3-5-S`)~`ADR2_J3-5-S`,
                             T~ADR2_JS))%>%
  rename(ADR2_J3_J5=`ADR2_J3-5`)%>%
  dplyr::select(ID,
                ADR2_J0,
                ADR2_J3_J5,
                ADR2_JS)

DATA <- read_delim("DATA/ADRECMO_DATA.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)%>%
  mutate(nom = str_to_upper(nom),
         prenom =str_to_upper(prenom),
         ID = paste0(numero,"-",nom,prenom))
        
df_list <- lst(DATA,ADR1,ADR2)

df <- reduce(df_list, left_join, by="ID")%>%
  filter(ID!="12-NANA")
```

# DATA Managment

```{r,include=FALSE}

df <- df %>%
  mutate(
    Temp_D0 = round(Temp_D0, 1),
    Age = as.numeric(Age),
    Gender = case_when(
      Gender == 0 ~ "Men",
      Gender == 1 ~ "Women",
      TRUE ~ as.character(Gender)
    ),
    Outcome = case_when(
      Outcome == 1 ~ "ECMO Weaning",
      Outcome == 2 ~ "Bridge to Transplant",
      Outcome == 3 ~ "Bridge to LVAD",
      Outcome == 4 ~ "Death",
      TRUE ~ as.character(Outcome)
    ),
    Pulsepressure_D0 = PAS_D0 - PAD_D0,
    icu_admiss_date = dmy(icu_admiss_date),
    icu_discharge_date = dmy(icu_discharge_date),
    date_levo = dmy(date_levo),
    ecmo_start_date = dmy(ecmo_start_date),
    ecmo_stop_date = dmy(ecmo_stop_date),
    Time_hosp = as.numeric(icu_discharge_date - icu_admiss_date),
    Time_ecmo = as.numeric(ecmo_stop_date - ecmo_start_date)
  )

```

# Check for normality 

```{r,include=FALSE}

for (var in var.cont) {
  if (var %in% colnames(T1)) {
    print(
      ggplot(T1, aes_string(var)) +
        geom_histogram(bins = 15, fill = "blue", alpha = 0.5) +
        ggtitle(paste("Distribuzione di", var))
    )
  } else {
    message(paste("❌ Variabile", var, "non trovata in T1"))
  }
}

```

# TABLE 1: description of baseline characteristics of the population at inclusion

```{r,include=FALSE}

var.cat <- c("Gender", "HTN", "DM", "Immunodepression", "CKD", "Neurologic_deficit", "Chronic_respiratory_disease", "HF", "Cardiac_arrest_before_canul", "IABP_D0", "Intubation", "EER_D0", "Outcome", "Betablocker", "Alphablocker", "Levosimendan", "Death")

var.cont <- c("Age", "BMI", "PAS_D0", "PAD_D0", "PAM_D0", "Pulsepressure_D0", "HR_D0", "Temp_D0", "Preecmo_tte_ef", "Preecmo_tte_vtiao", "Preecmo_tte_diam_LV", "Preecmo_tte_tapse", "Preecmo_tte_vmax_s_mit_lat_LV", "NADcum_D0", "DOBUcum_D0", "pH_D0", "Sao2_D0", "Paco2_D0", "Lact_D0", "Hco3_D0", "Creat_D0", "Urea_D0", "ALAT_D0", "ASAT_D0", "Bili_D0", "PT_D0", "Tropo_i_hs_D0", "Ntprobnp_D0", "DPP3_D0", "Plq_D0","ADRtot_D0","ADRtot_D3", "ADRtot_w", "Lenght_eer", "Lenght_vm", "Time_hosp", "Time_ecmo")

var.tot <- c(var.cat, var.cont)

T1 <- df %>%
  dplyr::select(all_of(var.tot))%>%
  mutate(across(all_of(var.cont), as.numeric), 
         across(all_of(var.cat), as.factor))


## Table 1 standard

table1 <- CreateTableOne(vars = var.tot, factorVars = var.cat, data =T1) # sans groupe de strata

table1 <- print(table1, nonnormal =var.cont)

table1_df <- as.data.frame(print(table1, nonnormal = var.cont, showAllLevels = TRUE))
table1_df$Variables <- row.names(table1_df)

row.names(table1_df)<-NULL
table1_df <- table1_df[, c(2,1)]

T1_n <- T1%>%
  summarise(across(everything(), ~ sum(is.na(.))))%>%
  pivot_longer(cols= everything(),
    values_to = "n_missing",
    names_to = "ID")%>%
  mutate(n_missing = paste0(n_missing,"/50"),
         ID = trimws(ID))

table1_df1 <- table1_df %>%
  mutate(ID = sub("\\(median \\[IQR\\]\\)|= 1 \\(\\%\\)|= Women \\(\\%\\)|\\(\\%\\)", "",Variables),
         ID = trimws(ID))

table_1_vf <- left_join(table1_df1,T1_n, by="ID")%>%
  select(Variables,n_missing,Overall)%>%
  mutate(n_missing = replace_na(n_missing,""))

table_1_vf <- left_join(table1_df1, T1_n, by = "ID") %>%
  mutate(N = 50 - as.numeric(sub("/50", "", replace_na(n_missing, "0")))) %>%  # Calcola N
  select(Variables, N, Overall)  

print(table_1_vf)

```

#Figure 1: Association between ADRB expressions and outcomes

```{r,include=FALSE}

##With composite outcome

F1 <- df %>%
  select(ID,ADR1_J0, ADR1_J3_J5, ADR1_JS, Outcome)%>%
  mutate(Outcomes = factor(case_when(Outcome=="Bridge to LVAD"|Outcome=="Bridge to Transplant"~ "Bridge to transplant or LVAD",T~Outcome), levels=c("ECMO Weaning","Bridge to transplant or LVAD","Death")))%>%
  pivot_longer(
    cols = c(ADR1_J0, ADR1_J3_J5, ADR1_JS), 
    names_to = "ADR1_time", 
    values_to = "value"
  )


count_data <- F1 %>%
  filter(!is.na(value))%>%
  group_by(ADR1_time, Outcomes) %>%
  summarise(n = n(), .groups = "drop")%>%
  filter(!is.na(Outcomes))

#Boxplot

Fig_1 <- ggplot(F1, aes(x = ADR1_time, y = value, fill = Outcomes)) +
  geom_boxplot(position = position_dodge(0.7), width = 0.6) +  
  scale_y_log10() +  
  labs(
    title = "Association between ADRB1 and outcomes",
    x = "times of measurement",
    y = "Percentage of T4 lymphocyte expressing ADRB1",
  ) +
  geom_text(data = count_data, aes(x = ADR1_time, y = min(F1$value, na.rm = TRUE) * 0.8, 
                                   label = paste0("n=", n), group = Outcomes), 
            position = position_dodge(0.7), size = 4, vjust = 1)+
    scale_x_discrete(labels = c("ADR1_J0" = "implantation", "ADR1_J3_J5" = "day 3 to 5", "ADR1_JS" = "explantation")) +
  theme(legend.position = "none")

print(Fig_1)
ggsave("Figures/boxplot_ADR1_Outcome.pdf", width = 8, height = 6)

## ADRB2 (J0, J3-5, JS) ~ death

F1_1 <- df %>%
  select(ID,ADR2_J0, ADR2_J3_J5, ADR2_JS, Outcome)%>%
  mutate(Outcomes = factor(case_when(Outcome=="Bridge to LVAD"|Outcome=="Bridge to Transplant"~ "Bridge to transplant or LVAD",T~Outcome), levels=c("ECMO Weaning","Bridge to transplant or LVAD","Death")))%>%
  pivot_longer(
    cols = c(ADR2_J0, ADR2_J3_J5, ADR2_JS), 
    names_to = "ADR2_time", 
    values_to = "value"
  )

count_data <- F1_1 %>%
  filter(!is.na(value))%>%
  group_by(ADR2_time, Outcomes) %>%
  summarise(n = n(), .groups = "drop")%>%
  filter(!is.na(Outcomes))

#Boxplot

Fig_1.1 <- ggplot(F1_1, aes(x = ADR2_time, y = value, fill = Outcomes)) +
  geom_boxplot(position = position_dodge(0.7), width = 0.6) +  
  scale_y_log10() +  
  labs(
    title = "Association between ADRB2 and outcomes",
    x = "times of measurement",
    y = "Percentage of T4 lymphocyte expressing ADRB2",
  ) +
  geom_text(data = count_data, aes(x = ADR2_time, y = min(F1_1$value, na.rm = TRUE) * 0.8, 
                                   label = paste0("n=", n), group = Outcomes), 
            position = position_dodge(0.7), size = 4, vjust = 1)+
      scale_x_discrete(labels = c("ADR1_J0" = "implantation", "ADR1_J3_J5" = "day 3 to 5", "ADR1_JS" = "explantation"))

print(Fig_1.1)
ggsave("Figures/boxplot_ADR2_Outcome.pdf", width = 8, height = 6)

Fig_1 <- Fig_1 + theme(plot.title = element_blank())
Fig_1.1 <- Fig_1.1 + theme(plot.title = element_blank())

Fig1_tot <- Fig_1 + Fig_1.1 +
  plot_annotation(title = "Association between percentage of T4 lymphocyte expressing ADRB and Outcome")
print(Fig1_tot)
ggsave("Figures/boxplot_ADR_Outcome_full.pdf", width = 12, height = 6)

```

#Figure 2: Association between VTI and ADRB1

```{r,include=FALSE}

colnames(df) <- gsub(",", ".", colnames(df))  


# VERSION AVEC ITV SEVRAGE

#ADRB1

F2_1 <- df %>%
  dplyr::select(jsev_epr_sev_t1_1.5l_itv,ADR1_J0,ADR1_J3_J5, ADR1_JS)%>%
  mutate(VTI_cut= case_when(jsev_epr_sev_t1_1.5l_itv>12~ ">12 cm",
               jsev_epr_sev_t1_1.5l_itv<=12~ "<=12 cm"))%>%
   filter(!is.na(ADR1_JS) & !is.na(jsev_epr_sev_t1_1.5l_itv))


#boxplot
Fig_2 <- ggplot(F2_1, aes(x= VTI_cut, y = ADR1_JS, fill = factor(VTI_cut))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between VTI at weaning and percentage of T4 lymphocyte expressing ADRB1", x = "VTI measured at 1.5L/min", y = "ADR1 measured \n the last day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_2)
ggsave("Figures/boxplot_VTI_ADR1.pdf", width = 8, height = 6)

#ADRB2

F2_2 <- df %>%
  dplyr::select(jsev_epr_sev_t1_1.5l_itv,ADR2_J0,ADR2_J3_J5, ADR2_JS)%>%
  mutate(VTI_cut= case_when(jsev_epr_sev_t1_1.5l_itv>12~ ">12 cm",
               jsev_epr_sev_t1_1.5l_itv<=12~ "<=12 cm"))%>%
   filter(!is.na(ADR2_JS) & !is.na(jsev_epr_sev_t1_1.5l_itv))


#boxplot
Fig_2.1 <- ggplot(F2_2, aes(x= VTI_cut, y = ADR2_JS, fill = factor(VTI_cut))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between VTI at weaning and percentage of T4 lymphocyte expressing ADRB2", x = "VTI measured at 1.5L/min", y = "ADR2 measured \n the last day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_2.1)
ggsave("Figures/boxplot_VTI_ADR2.pdf", width = 8, height = 6)



Fig_2 <- Fig_2 + theme(plot.title = element_blank())
Fig_2.1 <- Fig_2.1 + theme(plot.title = element_blank())

Fig2_tot <- Fig_2 + Fig_2.1 +
  plot_annotation(title = "Association between VTI at weaning and percentage of T4 lymphocyte expressing ADRB")
print(Fig2_tot)

ggsave("Figures/boxplot_VTI_ADR_1_2.pdf", width = 8, height = 6)

```


# Figure 3: Survival ~ ADR

```{r,include=FALSE}

#ADRB1
F3 <-df %>%
  dplyr::select(Outcome,outcome_date,icu_admiss_date, j28_deces,j28_date,ADR1_D0)%>%
  mutate(Outcome_censored = as.numeric(case_when(Outcome=="Death"~ "1",Outcome=="ECMO Weaning"|Outcome=="Bridge to Transplant"|Outcome=="Bridge to LVAD"~"0")),
         outcome_date = str_replace_all(outcome_date, "\\.", "/"),
         icu_admiss_date = str_replace_all(icu_admiss_date, "\\-", "/"),
         outcome_date= dmy(outcome_date),
         icu_admiss_date=ymd(icu_admiss_date),
         diff_days = as.numeric(outcome_date - icu_admiss_date),
         diff_days_J28 = case_when(diff_days>28 ~ 28,
                                   T~ diff_days),
         Outcome_censored_28 = case_when(diff_days>28 & Outcome_censored == 1~ 0,
                                         T~ Outcome_censored),
         ADR1_D0_median = case_when(ADR1_D0 > median(ADR1_D0, na.rm=T)~"High",
                                    ADR1_D0 <= median(ADR1_D0, na.rm=T)~"Low"))


km_fit <- survfit(Surv(diff_days_J28, Outcome_censored_28) ~ ADR1_D0_median, data = F3)

Fig_3 <- ggsurvplot(km_fit, data = F3, 
  pval = TRUE,               
  pval.coord = c(10, 0.1),   
  conf.int = FALSE,           
  risk.table = TRUE,         
  risk.table.height = 0.2,   
  risk.table.y.text = TRUE,  # 🔹 Rimuove "Strata" dalla tabella
  risk.table.title = "Number at Risk (number censored)",  # 🔹 Cambia il titolo della tabella
  xlim = c(0, 28),         # 🔹 Tronca l'asse X a 28 giorni
  break.time.by = 7,
  legend.title = " ",       # 🔹 Rimuove "Strata" dalla legenda
  palette = c("#E63946", "#457B9D"),
  ggtheme = theme_minimal()
)

Fig_3 <- Fig_3 + ggtitle("28-day Survival Curves for ADR1")
print(Fig_3)
Fig_3_combined <- plot_grid(Fig_3$plot, Fig_3$table, ncol = 1, align = "v", rel_heights = c(3, 1))

ggsave("Figures/Kaplan_Meier_plot_full.pdf", Fig_3_combined, width = 8, height = 6)

#ADRB2

F3_1 <-df %>%
  dplyr::select(Outcome,outcome_date,icu_admiss_date, j28_deces,j28_date,ADR2_D0)%>%
  mutate(Outcome_censored = as.numeric(case_when(Outcome=="Death"~ "1",Outcome=="ECMO Weaning"|Outcome=="Bridge to Transplant"|Outcome=="Bridge to LVAD"~"0")),
         outcome_date = str_replace_all(outcome_date, "\\.", "/"),
         icu_admiss_date = str_replace_all(icu_admiss_date, "\\-", "/"),
         outcome_date= dmy(outcome_date),
         icu_admiss_date=ymd(icu_admiss_date),
         diff_days = as.numeric(outcome_date - icu_admiss_date),
         diff_days_J28 = case_when(diff_days>28 ~ 28,
                                   T~ diff_days),
         Outcome_censored_28 = case_when(diff_days>28 & Outcome_censored == 1~ 0,
                                         T~ Outcome_censored),
         ADR2_D0_median = case_when(ADR2_D0 > median(ADR2_D0, na.rm=T)~"High",
                                    ADR2_D0 <= median(ADR2_D0, na.rm=T)~"Low"))


km_fit <- survfit(Surv(diff_days_J28, Outcome_censored_28) ~ ADR2_D0_median, data = F3_1)

Fig_3.1 <- ggsurvplot(km_fit, data = F3_1, 
  pval = TRUE,               
  pval.coord = c(10, 0.1),   
  conf.int = FALSE,           
  risk.table = TRUE,         
  risk.table.height = 0.2,   
  risk.table.y.text = TRUE,  # 🔹 Rimuove "Strata" dalla tabella
  risk.table.title = "Number at Risk (number censored)",  # 🔹 Cambia il titolo della tabella
  xlim = c(0, 28),         # 🔹 Tronca l'asse X a 28 giorni
  break.time.by = 7,
  legend.title = " ",       # 🔹 Rimuove "Strata" dalla legenda
  palette = c("#E63946", "#457B9D"),  
  ggtheme = theme_minimal()
)

Fig_3.1 <- Fig_3.1 + ggtitle("28-day Survival Curves for ADR2")
print(Fig_3.1)
Fig_3.1_combined <- plot_grid(Fig_3.1$plot, Fig_3.1$table, ncol = 1, align = "v", rel_heights = c(3, 1))

ggsave("Figures/Kaplan_Meier_plot_full.pdf", Fig_3.1_combined, width = 8, height = 6)


Fig_3_combined <- Fig_3_combined + theme(plot.title = element_blank())
Fig_3.1_combined <- Fig_3.1_combined + theme(plot.title = element_blank())
Fig3_tot <- Fig_3_combined + Fig_3.1_combined +
  plot_annotation(title = "28-day Survival Curves for ADR")
print(Fig3_tot)

ggsave("Figures/Kaplan_Meier_plot_full.pdf", Fig3_tot, width = 12, height = 6)

```

# Figure 4: ADR ~ beta blockers

```{r,include=FALSE}

#ADRB1

F4 <- df %>%
  dplyr::select(Betablocker,ADR1_J0,ADR1_J3_J5, ADR1_JS)%>%
  mutate(Betablocker= case_when(Betablocker==1~ "YES",
               Betablocker==0~ "NO"))%>%
   filter(!is.na(ADR1_J0) & !is.na(Betablocker))


#boxplot
Fig_4 <- ggplot(F4, aes(x= Betablocker, y = ADR1_J0, fill = factor(Betablocker))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Betablocker and ADRB1 at baseline", x = "Betablocker", y = "ADR1 measured at day 0 of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_4)
ggsave("Figures/boxplot_ADR1_BB.pdf", width = 8, height = 6)

#ADRB2

F4_1 <- df %>%
  dplyr::select(Betablocker,ADR2_J0,ADR2_J3_J5, ADR2_JS)%>%
  mutate(Betablocker= case_when(Betablocker==1~ "YES",
               Betablocker==0~ "NO"))%>%
   filter(!is.na(ADR2_J0) & !is.na(Betablocker))


#boxplot
Fig_4.1 <- ggplot(F4_1, aes(x= Betablocker, y = ADR2_J0, fill = factor(Betablocker))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Betablocker and ADRB2 at baseline", x = "Betablocker", y = "ADR2 measured at day 0 of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_4.1)
ggsave("Figures/boxplot_ADR2_BB.pdf", width = 8, height = 6)



Fig_4 <- Fig_4 + theme(plot.title = element_blank())
Fig_4.1 <- Fig_4.1 + theme(plot.title = element_blank())

Fig4_tot <- Fig_4 + Fig_4.1 +
  plot_annotation(title = "Association between BB at weaning and ARB")
print(Fig4_tot)

ggsave("Figures/boxplot_ADR_BB_full.pdf", width = 8, height = 6)

```

# Figure 5: ADR ~ levosimendan

```{r,include=FALSE}

#how many patients recevied Levosimendan after ICU admission

df <- df %>%
  mutate(
    levo_in_icu = case_when(
      Levosimendan == 1 & date_levo >= icu_admiss_date ~ "Yes",  
      Levosimendan == 1 & date_levo < icu_admiss_date ~ "No",
      Levosimendan == 0 ~ "Never"
    )
  )

table(df$levo_in_icu)

# 27 Patient received Levosimendan after ICU admission
# 2 Patient received Levosimendan before ICU admission
# 20 Patient never received Levosimendan

#Choise of ADRB1 at weaning

F5 <- df %>%
  dplyr::select(Levosimendan,ADR1_J0,ADR1_J3_J5, ADR1_JS)%>%
  mutate(Levosimendan= case_when(Levosimendan==1~ "YES",
               Levosimendan==0~ "NO"))%>%
   filter(!is.na(ADR1_JS) & !is.na(Levosimendan))


#boxplot
Fig_5 <- ggplot(F5, aes(x= Levosimendan, y = ADR1_JS, fill = factor(Levosimendan))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Levosimendan and ADRB1 at weaning", x = "Levosimendan", y = "ADR1 measured at last day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_5)
ggsave("Figures/boxplot_ADR1_Levosimendan.pdf", width = 8, height = 6)

#ADRB2

F5.1 <- df %>%
  dplyr::select(Levosimendan,ADR2_J0,ADR2_J3_J5, ADR2_JS)%>%
  mutate(Levosimendan= case_when(Levosimendan==1~ "YES",
               Levosimendan==0~ "NO"))%>%
   filter(!is.na(ADR2_JS) & !is.na(Levosimendan))


#boxplot
Fig_5.1 <- ggplot(F5.1, aes(x= Levosimendan, y = ADR2_JS, fill = factor(Levosimendan))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Levosimendan and ADRB2 at weaning", x = "Levosimendan", y = "ADR2 measured at last day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_5.1)
ggsave("Figures/boxplot_ADR2_Levosimendan.pdf", width = 8, height = 6)



Fig_5 <- Fig_5 + theme(plot.title = element_blank())
Fig_5.1 <- Fig_5.1 + theme(plot.title = element_blank())

Fig5_tot <- Fig_5 + Fig_5.1 +
  plot_annotation(title = "Association between Levosimendan and ARB at weaning")
print(Fig5_tot)

ggsave("Figures/boxplot_ADR_Levosimendan_full.pdf", width = 8, height = 6)

```

# Figure 5: ADR ~ dobutamin

```{r,include=FALSE}

F6 <- df %>%
  dplyr::select(dose_tot_dobu,ADR1_J0,ADR1_J3_J5, ADR1_JS)%>%
  mutate(dose_tot_dobu= case_when(dose_tot_dobu>0~ "YES",
               dose_tot_dobu==0~ "NO"))%>%
   filter(!is.na(ADR1_J0) & !is.na(dose_tot_dobu))


#boxplot
Fig_6 <- ggplot(F6, aes(x= dose_tot_dobu, y = ADR1_J0, fill = factor(dose_tot_dobu))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Dobutamin and ADRB1 at baseline", x = "Dobutamin", y = "ADR1 measured at day 0 of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_6)
ggsave("Figures/boxplot_ADR1_Dobu.pdf", width = 8, height = 6)

#ADRB2

F6.1 <- df %>%
  dplyr::select(dose_tot_dobu,ADR2_J0,ADR2_J3_J5, ADR2_JS)%>%
  mutate(dose_tot_dobu= case_when(dose_tot_dobu>0~ "YES",
               dose_tot_dobu==0~ "NO"))%>%
   filter(!is.na(ADR2_J0) & !is.na(dose_tot_dobu))


#boxplot
Fig_6.1 <- ggplot(F6.1, aes(x= dose_tot_dobu, y = ADR2_J0, fill = factor(dose_tot_dobu))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Dobutamin and ADRB2 at baseline", x = "Dobutamin", y = "ADR2 measured at day 0 of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_6.1)
ggsave("Figures/boxplot_ADR2_Dobutamin.pdf", width = 8, height = 6)



Fig_6 <- Fig_6 + theme(plot.title = element_blank())
Fig_6.1 <- Fig_6.1 + theme(plot.title = element_blank())

Fig6_tot <- Fig_6 + Fig_6.1 +
  plot_annotation(title = "Association between Dobutamin and ARB at Day 0 of ECMO")
print(Fig6_tot)

ggsave("Figures/boxplot_ADRbase_Dobutamin_full.pdf", width = 8, height = 6)

## Dobu at JS

#ADR1

F6.2 <- df %>%
  dplyr::select(dose_tot_dobu,ADR1_J0,ADR1_J3_J5, ADR1_JS)%>%
  mutate(dose_tot_dobu= case_when(dose_tot_dobu>0~ "YES",
               dose_tot_dobu==0~ "NO"))%>%
   filter(!is.na(ADR1_JS) & !is.na(dose_tot_dobu))


#boxplot
Fig_6.2.2 <- ggplot(F6.2, aes(x= dose_tot_dobu, y = ADR1_JS, fill = factor(dose_tot_dobu))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Dobutamin and ADRB1 at weaning", x = "Dobutamin", y = "ADR1 measured at dlast day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_6.2.2)
ggsave("Figures/boxplot_ADR1_Dobu_weaning.pdf", width = 8, height = 6)

#ADRB2

F6.3 <- df %>%
  dplyr::select(dose_tot_dobu,ADR2_J0,ADR2_J3_J5, ADR2_JS)%>%
  mutate(dose_tot_dobu= case_when(dose_tot_dobu>0~ "YES",
               dose_tot_dobu==0~ "NO"))%>%
   filter(!is.na(ADR2_JS) & !is.na(dose_tot_dobu))


#boxplot
Fig_6.3 <- ggplot(F6.3, aes(x= dose_tot_dobu, y = ADR2_JS, fill = factor(dose_tot_dobu))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Dobutamin and ADRB2 at weaning", x = "Dobutamin", y = "ADR2 measured at last day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_6.3)
ggsave("Figures/boxplot_ADR2_Dobutamin_weaning.pdf", width = 8, height = 6)



Fig_6.2.2 <- Fig_6.2.2 + theme(plot.title = element_blank())
Fig_6.3 <- Fig_6.3 + theme(plot.title = element_blank())

Fig6_tot_weaning <- Fig_6.2.2 + Fig_6.3 +
  plot_annotation(title = "Association between Dobutamin and ARB at last Day of ECMO")
print(Fig6_tot_weaning)

ggsave("Figures/boxplot_ADR_Dobutamin_full.pdf", width = 8, height = 6)

## ADR ~ cumulative dose dobu (median)

#ADRB1

F6.4 <- df %>%
  dplyr::select(dose_tot_dobu,ADR1_J0,ADR1_J3_J5, ADR1_JS)%>%
  mutate(dobu_cut= case_when(dose_tot_dobu> median(ADR1_JS, na.rm=T)~"High",
               dose_tot_dobu< median(ADR1_JS, na.rm=T)~"Low"))%>%
   filter(!is.na(ADR1_JS) & !is.na(dose_tot_dobu))


#boxplot
Fig_6.4 <- ggplot(F6.4, aes(x= dobu_cut, y = ADR1_JS, fill = factor(dobu_cut))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between cumulative dose of dobutamina and ADRB1 at weaning", x = "dobutamin", y = "ADR1 measured the last day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_6.4)
ggsave("Figures/boxplot_cumdobu_ADR1.pdf", width = 8, height = 6)

#ADRB2

F6.5 <- df %>%
  dplyr::select(dose_tot_dobu,ADR2_J0,ADR2_J3_J5, ADR2_JS)%>%
  mutate(dobu_cut= case_when(dose_tot_dobu> median(ADR2_JS, na.rm=T)~"High",
               dose_tot_dobu< median(ADR2_JS, na.rm=T)~"Low"))%>%
   filter(!is.na(ADR2_JS) & !is.na(dose_tot_dobu))


#boxplot
Fig_6.5 <- ggplot(F6.5, aes(x= dobu_cut, y = ADR2_JS, fill = factor(dobu_cut))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between cumulative dose of dobutamina and ADRB2 at weaning", x = "dobutamin", y = "ADR2 measured the last day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_6.5)
ggsave("Figures/boxplot_cumdobu_ADR2.pdf", width = 8, height = 6)


Fig_6.4 <- Fig_6.4 + theme(plot.title = element_blank())
Fig_6.5 <- Fig_6.5 + theme(plot.title = element_blank())

Fig6_tot_dobucum_weaning <- Fig_6.4 + Fig_6.5 +
  plot_annotation(title = "Association between cumulative dose of Dobutamin and ARB at last Day of ECMO")
print(Fig6_tot_dobucum_weaning)

ggsave("Figures/boxplot_ADR_cum_Dobutamin_full.pdf", width = 8, height = 6)

```

# Figure 6: ADR ~ noradrenalin

```{r,include=FALSE}

#ADR1

F7 <- df %>%
  dplyr::select(dose_tot_NAd,ADR1_J0,ADR1_J3_J5, ADR1_JS)%>%
  mutate(dose_tot_NAd= case_when(dose_tot_NAd>0~ "YES",
               dose_tot_NAd==0~ "NO"))%>%
   filter(!is.na(ADR1_JS) & !is.na(dose_tot_NAd))


#boxplot
Fig_7 <- ggplot(F7, aes(x= dose_tot_NAd, y = ADR1_JS, fill = factor(dose_tot_NAd))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Noradrenalin and ADRB1 at weaning", x = "Noradrenalin", y = "ADR1 measured at last day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_7)
ggsave("Figures/boxplot_ADR1_Nora_weaning.pdf", width = 8, height = 6)

#ADRB2

F7.1 <- df %>%
  dplyr::select(dose_tot_NAd,ADR2_J0,ADR2_J3_J5, ADR2_JS)%>%
  mutate(dose_tot_NAd= case_when(dose_tot_NAd>0~ "YES",
               dose_tot_NAd==0~ "NO"))%>%
   filter(!is.na(ADR2_JS) & !is.na(dose_tot_NAd))


#boxplot
Fig_7.1 <- ggplot(F7.1, aes(x= dose_tot_NAd, y = ADR2_JS, fill = factor(dose_tot_NAd))) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7) +
  scale_fill_manual(values = c("white", "gray70")) +
  labs(title = "Association between Noradrenalin and ADRB2 at weaning", x = "Noradrenalin", y = "ADR2 measured at last day of ECMO") +
  theme_minimal() +
  theme(legend.position = "none")  

print(Fig_6.3)
ggsave("Figures/boxplot_ADR2_Nora_weaning.pdf", width = 8, height = 6)



Fig_7 <- Fig_7 + theme(plot.title = element_blank())
Fig_7.1 <- Fig_7.1 + theme(plot.title = element_blank())

Fig7_tot <- Fig_7 + Fig_7.1 +
  plot_annotation(title = "Association between Noradrenalin and ARB at last Day of ECMO")
print(Fig7_tot)

ggsave("Figures/boxplot_ADR_Nora_full.pdf", width = 8, height = 6)

```