---
title: "ADRECMO"
author: "Jolie Bruno, Antoine Kimmoun, Kevin Duarte, Feriel Azibani, Benjamin Deniaun"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
    latex_engine: pdflatex
  word_document: null
editor_options:
  chunk_output_type: console
always_allow_html: yes
---

```{=html}
<style type="text/css">

h1.title {
  font-size: 38px;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  text-align: center;
}
</style>
```

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
# markdown options depending on the output (html with R code and Word without)
output <- knitr::opts_knit$get("rmarkdown.pandoc.to")
if (output=="html") knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo = TRUE, comment = "", fig.align = "center", fig.width= 17, fig.asp =0.70, out.width = "90%")
if (output=="docx") knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo = FALSE, comment = "", fig.align = "center", fig.width= 17, fig.asp =0.70, out.width = "90%")
if (output=="pdf") knitr::opts_chunk$set(message=FALSE, warning=FALSE, echo = FALSE, comment = "", fig.align = "center", fig.width= 17, fig.asp =0.70, out.width = "90%")
options(max.print = .Machine$integer.max)
```

#Package
```{r,include=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(stringr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(gtsummary)
library(sjlabelled)
library(readxl)
library(purrr)
library(tableone)
library(naniar)
library(survival)
library(survminer)
library(patchwork)
library(ggpmisc)
library(cowplot)  
library(officer)
library(flextable)
library(lme4)
library(lmerTest)
library(emmeans)
library(forcats)
```

# Method

This study is a monocentric, prospective, observational investigation conducted in Nancy Hospital between 10.10.2017 and 24.05.2022, enrolling 50 adult patients with refractory cardiogenic shock requiring veno-arterial extracorporeal membrane oxygenation (VA-ECMO) support (NCT03327493).

The study was carried out in accordance with the Declaration of Helsinki and approved by the local institutional ethics committee. Written informed consent was obtained from all patients or their legal representatives prior to inclusion.

## Patient Population

Cardiogenic shock was defined as a systolic blood pressure (SBP) <90 mmHg or a mean arterial pressure (MAP) <65 mmHg despite adequate volume resuscitation, accompanied by signs of peripheral hypoperfusion (e.g., cold extremities, oliguria, altered mental status) and a cardiac index <2.2 L/min/m². Refractoriness was defined as a hyporesponsiveness to norepinephrine and/or the persistence of profound clinical signs of hypoperfusion despite optimal resuscitation efforts.
Eligible patients were adults (≥18 years old) affiliated with a  health insurance system. Cardiogenic shock etiologies included acute ischemic coronary artery disease, primary or ischemic-related dilated cardiomyopathy, myocarditis and stress-induced cardiomyopathy. Patients were excluded if cardiogenic shock was secondary to cardiotoxic poisoning, if they were pregnant, or were under legal protective supervision (e.g., court-appointed guardianship).

## Data Collection

Clinical, laboratory and echocardiographic parameters were prospectively collected at three predefined timepoints:Day 0 (ECMO implantation), Day 3–5 (early course) and Day of ECMO weaning (ECMO explantation). Blood samples were drawn at each timepoint, immediately centrifuged and plasma was stored at –80°C until analysis. Samples were then transferred to the XXX for cells analysis and INSERM UMR942 laboratory (Paris, France) for cytokine analysis. Quantification of adrenoreceptors β1 and β2 on monocytes and CD4⁺ T lymphocytes was performed using flow cytometry. Cytokine concentrations in plasma, including interleukin (IL) -33, IL-1β, IL-2, IL-4, IL-5, IL-6, IL-8/CXCL8, IL-10, GDF-15, IL-12p70, IL-17A/CTLA-8, IFN-γ, TNF-α and granzyme B, were measured using [XXX].

## Assessment of ADRB1 and ADRB2

The expression of β1 (ADRB1) and β2 (ADRB2) adrenergic receptors on circulating immune cell, namely CD4+ T lymphocytes and monocytes, was assessed via flow cytometry. XXX [to be completed from Manon]. Results were expressed as the percentage of cells that were positive for the respective expression receptors.

##Cytokine Quantification

The aforementioned panel of circulating cytokines and immune mediators was quantified from plasma samples. These biomarkers were selected for their roles in pro- and anti-inflammatory signaling. Measurements were performed using a multiplex bead-based immunoassay (Luminex platform), following the manufacturer’s instructions. XXX [to be completed from Malha].

##ECMO Weaning Protocol

A standardized weaning protocol was applied to all patients prior to ECMO explantation. At a constant mean arterial pressure (MAP) between 65 and 75 mmHg, cardiac output was assessed by transthoracic echocardiography at two predefined timepoints: 1. Baseline under full support (ECLS flow at 3.0 L/min) and 2. Low-flow test after 45 minutes of reduced ECMO flow (1.5 L/min).
Measurements included velocity time integral (VTI) at the left ventricular outflow tract and left ventricular ejection fraction. Changes in norepinephrine and dobutamin dosage during this period were also documented. The decision to explant ECMO was based on clinical tolerance, hemodynamic stability and echocardiographic parameters during the weaning trial.

## Statistical Analysis


Baseline clinical, demographic and hemodynamic characteristics were summarized for the overall study population (n = 50). One patient was excluded from the analysis due to being under legal guardianship. Categorical variables were reported as counts and percentages, while continuous variables were expressed as medians and interquartile ranges (IQR). The normality of continuous variables was assessed visually using histograms.

To explore the distribution of ADRB1 and ADRB2 expression over time in relation to clinical outcomes, boxplots were generated at three timepoints (implantation, day 3–5, and explantation). To evaluate the temporal dynamics of ADRB1 and ADRB2 expression across clinical outcomes, we used linear mixed-effects models with time, outcome group (ECMO weaning, bridge to LVAD/transplant, or death), and their interaction as fixed effects. Random intercepts and slopes for time were included to account for intra-patient variability. When appropriate, receptor expression values were log10-transformed. Interaction terms were tested using likelihood ratio tests between full and reduced models, and Type III ANOVA with Kenward-Roger approximation was used for fixed effects.


To assess the association between ADRB1 and ADRB2 expression at the time of ECMO weaning and left ventricular outflow tract velocity-time integral (VTI) measured at 1.5 L/min ECMO flow, patients were stratified according to a VTI threshold of 12 cm. Mann–Whitney U tests were used to compare log-transformed ADRB expression between the two groups (≤12 cm vs >12 cm). Results were visualized using boxplots.

#FONCTION

```{r,include=FALSE}
source("FUNCTIONS/Examples - RCS effect + interaction cont x factor + check normality 181224.R")
source("FUNCTIONS/Plot_cytokine.R")

```

#DATA LOAD

```{r,include=FALSE}
#ADR LYMPHOCYTES
fichier <- "DATA/df_i_stabilized.xlsx"
feuilles <- excel_sheets(fichier)
data_list <- lapply(feuilles, function(feuille) {
  read_excel(fichier, sheet = feuille)
})


ADR1_lympho <- data_list[[1]] %>%
  select( "Patient","Jour",`Paramètre (Gate)`, `%Gated`)%>%
  rename(value = `%Gated`,
         parametre =`Paramètre (Gate)`)%>%
  mutate(parametre = case_when(parametre == "CD3+/CD4+" ~ "T4",
                               
                              parametre == "ADR1 (CD3+/CD4+)" ~ "ADR1L",
                               
                              parametre == "INFg+ (CD3+ CD4+)" ~ "L_TH1",
                              parametre == "IL4+ (CD3+ CD4+)" ~ "L_TH2",
                              parametre == "IL4+ INFg+ (CD3+ CD4+)" ~ "mixte",
                              
                              parametre == "ADR1 (INFg+)" ~ "ADR1L_TH1",
                              parametre == "ADR1 (IL4+)" ~ 
"ADR1L_TH2",
                              parametre == "ADR1 (IL4+ INFg+)" ~ "ADR1L_mixte",
                              
                              parametre == "IL17+ (CD3+ CD4+)" ~ "L_IL17",
                              parametre == "IL17+ INFg+ (CD3+ CD4+)" ~ "L_IL17_pathogen",
                              
                              parametre == "ADR1 (IL17+)" ~ "L_ADR1L_IL17",
                              parametre == "ADR1 (IL17+ INFg+)" ~ "L_ADR1L_IL17_pathogen",
                              
                               parametre == "IL17+ INFg+ (CD3+ CD4+)" ~ "L_IL17_pathogen",
                               
                               T~parametre),
         ID =str_replace(Patient, "Patient (\\d+) (\\w+)", "\\1-\\2"),
         value = as.numeric(value))%>%
  pivot_wider(
    names_from = c("parametre","Jour"),
    values_from = "value",
    values_fn = mean,  
    values_fill = NA
  )%>%
    filter(ID!="12-AA")

#CHECK AK OK


ADR2_lympho <- data_list[[2]] %>%
  select( "Patient","Jour",Gate, `%Gated`)%>%
  rename(value = `%Gated`,
         parametre = Gate)%>%
  mutate(parametre = case_when(
                              parametre == "ADR2 (CD3+/CD4+)" ~ "ADR2L",
                               
                            
                              parametre == "ADR2 (INFg+)" ~ "ADR2L_TH1",
                              parametre == "ADR2 (IL4+)" ~ "ADR2L_TH2",
                              parametre == "ADR2 (IL4+ INFg+)" ~ "ADR2L_mixte",
                              
                              
                              parametre == "ADR2 (IL17+)" ~ "L_ADR2L_IL17",
                              parametre == "ADR2 (IL17+ INFg+)" ~ "L_ADR2L_IL17_pathogen",
                              
                               
                               T~parametre),
         ID =str_replace(Patient, "Patient (\\d+) (\\w+)", "\\1-\\2"),
         value = as.numeric(value))%>%
  pivot_wider(
    names_from = c("parametre","Jour"),
    values_from = "value",
    values_fn = mean,  
    values_fill = NA
  )%>%
    filter(ID!="12-AA")


#CHECK AK OK

##Add Monocytes

fichier_mono <- "DATA/monocytes.xlsx"
feuilles_mono <- excel_sheets(fichier_mono)
data_mono <- lapply(feuilles_mono, function(feuille_mono) {
  read_excel(fichier_mono, sheet = feuille_mono)
})

ADR1_mono <- data_mono[[1]] %>%
  select("Patient", "Jour","Paramètre (Gate)", "%Gated") %>%
  rename(value = `%Gated`,
         parametre = "Paramètre (Gate)") %>%
  mutate(parametre= case_when(parametre == "ADR1 (Monocytes)"~"M_ADR1",
                              parametre == "IL1+ (Monocytes)"~"M_IL1",
                              parametre == "IL10+ (Monocytes)"~"M_IL10",
                              parametre == "TNF+ (Monocytes)"~"M_TNF",
                              T~parametre                              ),
    # Create clean ID and normalize 'Jour'
    ID = str_replace(Patient, "Patient (\\d+) (\\w+)", "\\1-\\2")
  ) %>%
  select(ID, Jour,parametre,value )%>%
  pivot_wider(
    names_from = c("parametre","Jour"),
    values_from = "value",
    values_fn = mean,
    values_fill = NA
  ) %>%
    filter(ID!="12-AA")
#CHECK AK OK

ADR2_mono <- data_mono[[2]] %>%
  select("Patient", "Jour","Paramètre (Gate)", "%Gated") %>%
  rename(value = `%Gated`,
         parametre = "Paramètre (Gate)") %>%
mutate(parametre= case_when(parametre == "ADR2 (Monocytes)"~"M_ADR2",T~parametre),
    # Create clean ID and normalize 'Jour'
    ID = str_replace(Patient, "Patient (\\d+) (\\w+)", "\\1-\\2")
  )%>%
  filter(parametre=="M_ADR2")%>%
    select(ID, Jour,parametre,value )%>%
  # Reshape from long to wide format
  pivot_wider(
    names_from = c("parametre","Jour"),
    values_from = "value",
    values_fn = mean,
    values_fill = NA
  ) %>%
    filter(ID!="12-AA")
#CHECK AK OK

df_list_1 <- lst(ADR1_lympho, ADR2_lympho, ADR1_mono, ADR2_mono)

# COUNT VALID ID
count_valid_ID <- function(data, id_col = "ID") {
  # Fonction qui compte les ID non-NA dans un seul data.frame
  count_one <- function(df) {
    df %>%
      mutate(
        ID_num = as.integer(str_remove(.data[[id_col]], "-.*")),
        ID_CAT = if_else(!is.na(ID_num), 1L, 0L)
      ) %>%
      summarise(n_valid = sum(ID_CAT, na.rm = TRUE)) %>%
      pull(n_valid)
  }
    imap_dfr(
      data,
      ~ tibble(
          dataset  = .y,
          n_valid  = count_one(.x)
        ),
      .id = NULL
    )
  }

count_valid_ID(df_list_1)

df_1 <- reduce(df_list_1, full_join, by = "ID")%>%
  mutate(any_non_na = as.integer(
      rowSums(
        !is.na( select(., starts_with("ADR")) ),
        na.rm = TRUE       
      ) > 0                  # true if at least one non NA
    )
  )

#IDENTIFY WHICH ID NOT PRESENT IN ONE AND PRESENT IN OTHER
id_presence <- imap_dfr(df_list_1,           
  ~ tibble(ID      = .x$ID,   
           dataset = .y)) %>%  
  mutate(present = 1L) %>%     
  distinct() %>%               
  pivot_wider(
    names_from   = dataset, 
    values_from  = present,    
    values_fill  = list(present = 0L)
  )
diff_IDs <- id_presence %>%
  filter(if_any(-ID, ~ .x == 0))


DATA <- read_delim("DATA/ADRECMO_DATA.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)%>%
  mutate(nom = str_to_upper(nom),
         prenom =str_to_upper(prenom),
         ID = paste0(numero,"-",nom, prenom))%>%
    filter(ID != "12-NANA")

#CHECK AK OK
df_list_2 <- lst(df_1,DATA)

df_2 <- reduce(df_list_2, left_join, by = "ID")

#CHECK AK OK

# --- Constantes et chemins ---
fichier_cyto <- "DATA/cytokines.xlsx"
feuille_idx  <- 7  # on lit la 7e feuille

# --- Variables attendues ---
ck_vars <- c(
  "Sample ID number",
  "[IL-33] in pg/mL","[IL1b] in pg/mL","[IL2] in pg/mL","[IL4] in pg/mL",
  "[IL5] in pg/mL","[IL6] in pg/mL","[IL8/CXCL8]  in pg/mL","[IL10] in pg/mL",
  "[GDF-15] in pg/mL","[IL12p70] in pg/mL","[IL17A/CTLA-8] in pg/mL",
  "[IFNg] in pg/mL","[TNFa] in pg/mL","[Granzyme B]  in pg/mL"
)

# --- Fonctions utilitaires ---
normalize_jour <- function(sample_id) {
  raw <- sample_id %>%
    str_extract("J[A-Z0-9\\-\\.]+") %>%
    str_to_upper() %>%
    str_replace_all("-", "_") %>%
    str_replace_all("\\.", "")
  case_when(
    raw %in% c("J3","J4","J5","J3_5","J3_J5") ~ "J3_J5",
    raw %in% c("JS","JSEVRAGE")               ~ "JS",
    raw == "J0"                                ~ "J0",
    TRUE                                       ~ NA_character_
  )
}

clean_measure_names <- function(x) {
  x %>%
    str_remove_all("\\[|\\]") %>%
    str_remove_all("\\s+in\\s+pg/mL") %>%
    str_squish() %>%
    str_replace_all("\\s+", "")
}

to_numeric_safe <- function(x) {
  x %>%
    as.character() %>%
    str_replace_all(",", ".") %>%
    str_remove_all("\\*") %>%
    str_replace_all("OOR.*|<.*|undetectable|low|LOQ|ND", "0.0000001") %>%
    as.numeric()
}

# --- Vérifications d’entrée ---
all_cols <- names(read_excel(fichier_cyto, sheet = feuille_idx, n_max = 0))
stopifnot(all(ck_vars %in% all_cols))

# --- 1) Lecture de la feuille cytokines (7) ---
df_cyto_raw <- read_excel(fichier_cyto, sheet = feuille_idx) %>%
  select(all_of(ck_vars))

# --- 2) Préparation / nettoyage ---
df_ids <- df_2 %>%
  select(ID) %>%
  mutate(ID_num = str_extract(ID, "^\\d{1,2}") %>%
           as.integer() %>%
           as.character())

df_cyto_clean <- df_cyto_raw %>%
  rename(SampleID = `Sample ID number`) %>%
  mutate(
    ID_num = str_extract(SampleID, "\\d{2}") %>%
      as.integer() %>%
      as.character(),
    Jour   = normalize_jour(SampleID)
  ) %>%
  left_join(df_ids, by = "ID_num") %>%
  select(-ID_num) %>%
  rename_with(clean_measure_names, .cols = -c(SampleID, ID, Jour)) %>%
  mutate(across(-c(SampleID, ID, Jour), to_numeric_safe))

# --- 3) Reshape long -> wide ---
df_cyto_long <- df_cyto_clean %>%
  select(-SampleID) %>%
  filter(!is.na(ID), !is.na(Jour)) %>%
  pivot_longer(
    cols      = -c(ID, Jour),
    names_to  = "cytokine",
    values_to = "value"
  ) %>%
  mutate(var_name = paste0(cytokine, "_", Jour)) %>%
  select(ID, var_name, value) %>%
  pivot_wider(names_from = var_name, values_from = value)

# --- 4) Fusion finale ---
df <- df_2 %>%
  left_join(df_cyto_long, by = "ID")

# --- 5) Vérification optionnelle ---
verif <- read_excel("DATA/ADRECMO_Verif.xlsx") %>%
  mutate(
    Initiales = sub("-", "", Initiales),
    ID        = paste0(`N° patient`, "-", Initiales) %>%
      str_remove("^0+")
  ) %>%
  rename(cause = `Origine du choc cardiogénique`) %>%
  select(ID, cause, CA)

df <- df %>%
  left_join(verif, by = "ID")

# --- 6) Sous-table bio + mapping outcomes ---
df_bio <- df %>%
  select(
    ID, Outcome,
    starts_with("T4_"),
    starts_with("ADR1L_"),
    starts_with("ADR2L_"),
    starts_with("L_TH1_"),
    starts_with("L_TH2_"),
    starts_with("mixte_"),
    starts_with("bINFg"),
    starts_with("IL"),
    starts_with("IFNg"),
    starts_with("TNF"),
    starts_with("GDF-15"),
    starts_with("GranzymeB"),
    starts_with("M_"),
    starts_with("L_"),
    starts_with("mix")
  ) %>%
  mutate(
    Outcome = case_when(
      Outcome %in% c(2, 3) ~ "Bridge to Transplant or LVAD",
      Outcome == 1         ~ "ECMO Weaning",
      Outcome == 4         ~ "Death",
      TRUE                 ~ as.character(Outcome)
    )
  )
```

# TIMEPOINT CHECK OF THE BLOOD SAMPLE AT DAY 0

```{r,include=TRUE}
# Blood sample at day 0 before ECMO implant?

df %>%
  select(J0_date, j0_bilan_date) %>%
  distinct() %>%
  slice(1:15)

df <- df %>%
  mutate(
    J0_date = as.Date(J0_date, format = "%d.%m.%Y"),
    j0_bilan_date = as.Date(j0_bilan_date, format = "%d.%m.%Y"),
    bilan_after_ecmo = case_when(
      is.na(J0_date) | is.na(j0_bilan_date) ~ NA_character_,
      j0_bilan_date > J0_date ~ "Yes",
      TRUE ~ "No"
    )
  )

table(df$bilan_after_ecmo, useNA = "ifany")
#CHECK AK OK

df %>%
  filter(bilan_after_ecmo == "Yes") %>%
  select(ID, j0_bilan_date, J0_date, bilan_after_ecmo)

#CHECK AK OK


df_clean <- df %>%
  mutate(
    j0_bilan_date = as.Date(j0_bilan_date, format = "%d.%m.%Y"),
    J0_date = as.Date(J0_date, format = "%d.%m.%Y"),
    datetime_bilan = as.POSIXct(paste(j0_bilan_date, as.character(j0_bilan_heure)), format = "%Y-%m-%d %H:%M", tz = "UTC"),
    datetime_ecmo  = as.POSIXct(paste(J0_date, as.character(J0_heure)), format = "%Y-%m-%d %H:%M", tz = "UTC"),
    time_diff_mins = as.numeric(difftime(datetime_bilan, datetime_ecmo, units = "mins"))
  ) %>%
  filter(!is.na(time_diff_mins))


df_clean %>%
    mutate(time_diff_hours = time_diff_mins / 60) %>%
  summarise(
    mean_minutes = mean(time_diff_hours),
    sd_minutes = sd(time_diff_hours),
    min = min(time_diff_hours),
    max = max(time_diff_hours)
  )

```

Day-0 blood samples were collected on average 3.47 hours (SD 2.43 hours) before ECMO implantation, ranging from -11 hours before to the exact time of ECMO cannulation.

# DATA Management

```{r,include=TRUE}
df <- df %>%
  mutate(
    Temp_D0 = round(Temp_D0, 1),
    Age = as.numeric(Age),
    Gender = case_when(
      Gender == 0 ~ "Men",
      Gender == 1 ~ "Women",
      TRUE ~ as.character(Gender)
    ),
    
    Preecmo_tte_diam_LV = Preecmo_tte_diam_LV * 10,
    
    Outcome = case_when(
      Outcome == 1 ~ "ECMO Weaning",
      Outcome == 2 ~ "Bridge to Transplant",
      Outcome == 3 ~ "Bridge to LVAD",
      Outcome == 4 ~ "Death",
      TRUE ~ as.character(Outcome)
    ),

    Pulsepressure_D0 = PAS_D0 - PAD_D0,
    

    outcome_date = str_replace_all(outcome_date, "\\.", "/"),
    icu_admiss_date = str_replace_all(icu_admiss_date, "\\.", "/"),
    icu_discharge_date = str_replace_all(icu_discharge_date, "\\.", "/"),

    deces_date = dmy(deces_date),
    icu_admiss_date = dmy(icu_admiss_date),
    icu_discharge_date = dmy(icu_discharge_date),
    date_levo = dmy(date_levo),
    ecmo_start_date = dmy(ecmo_start_date),
    ecmo_stop_date = dmy(ecmo_stop_date),
    outcome_date = case_when(is.na(deces_date)~ icu_admiss_date + 90,
                             T~deces_date),

    diff_days = time_length(interval(icu_admiss_date, outcome_date), "days"),
    Time_hosp = time_length(interval(icu_admiss_date, icu_discharge_date), "days"),
    as.numeric(icu_discharge_date - icu_admiss_date),
    Time_ecmo = time_length(interval(ecmo_start_date, ecmo_stop_date), "days"),
    Time_levo = time_length(interval(ecmo_start_date, date_levo), "days"),
    jsev_date = dmy(jsev_date),
    Time_weaning =time_length(interval(ecmo_start_date, jsev_date), "days"),
    Levo_before_ecmo_end = case_when(
      Levosimendan == 1 &
        date_levo     >= ecmo_start_date &
        date_levo     <= ecmo_stop_date ~ 1L,
      TRUE ~ 0L),

    ADR1L_J0_median = case_when(
      ADR1L_J0 > median(ADR1L_J0, na.rm = TRUE) ~ "High",
      ADR1L_J0 <= median(ADR1L_J0, na.rm = TRUE) ~ "Low"
    ),
    ADR2L_J0_median = case_when(
      ADR2L_J0 > median(ADR2L_J0, na.rm = TRUE) ~ "High",
      ADR2L_J0 <= median(ADR2L_J0, na.rm = TRUE) ~ "Low"
    ),

    M_ADR1_J0_median = case_when(
      M_ADR1_J0 > median(M_ADR1_J0, na.rm = TRUE) ~ "High",
      M_ADR1_J0 <= median(M_ADR1_J0, na.rm = TRUE) ~ "Low",
      TRUE ~ NA_character_
    ),
    M_ADR2_D0_median = case_when(
      M_ADR2_J0 > median(M_ADR2_J0, na.rm = TRUE) ~ "High",
      M_ADR2_J0 <= median(M_ADR2_J0, na.rm = TRUE) ~ "Low",
      TRUE ~ NA_character_
    ),

    Betablocker = case_when(
      Betablocker == 1 ~ "YES",
      Betablocker == 0 ~ "NO"
    ),

    dose_tot_dobu_yes_no = droplevels(as.factor(case_when(
      dose_tot_dobu > 0 ~ "YES",
      dose_tot_dobu == 0 ~ "NO"
    ))),

    levo_in_icu = case_when(
      Levosimendan == 1 & Time_levo >= 0 ~ "Yes",  
      Levosimendan == 1 & Time_levo < 0 ~ "No",
      Levosimendan == 0 ~ "No"
    ),

    Levosimendan = case_when(
      Levosimendan == 1 ~ "YES",
      Levosimendan == 0 ~ "NO",
      TRUE ~ NA_character_ 
    ),
    ische = relevel(as.factor(cause), ref=c("Non ischémique")),
    mean_NAD_24H = NADcum_D0*1000/Weight/3600,
    mean_DOB_24H = DOBUcum_D0*1000/Weight/3600
  )

#CHECK AK OK
var.cat <- c("Gender", "HTN", "DM", "Immunodepression", "CKD", "Neurologic_deficit", "Chronic_respiratory_disease", "HF", "Cardiac_arrest_before_canul", "cause","ische", "IABP_D0", "Intubation", "EER_D0", "Outcome", "Betablocker", "Alphablocker", "Levosimendan", "j90_deces", "j28_deces")

var.cont <- c("Age", "BMI", "PAS_D0", "PAD_D0", "PAM_D0", "Pulsepressure_D0", "HR_D0", "Temp_D0", "Preecmo_tte_ef", "Preecmo_tte_vtiao", "mean_NAD_24H", "mean_DOB_24H", "pH_D0", "Sao2_D0", "Paco2_D0", "Lact_D0", "Hco3_D0", "Creat_D0", "Urea_D0", "ALAT_D0", "ASAT_D0", "Bili_D0", "PT_D0", "Tropo_i_hs_D0", "Ntprobnp_D0", "DPP3_D0", "Plq_D0", "Lenght_eer", "Lenght_vm", "Time_hosp", "Time_ecmo")

var.tot <- c(var.cat, var.cont)

```

# TABLE 1: Characteristics of the population

```{r,include=FALSE}

T1 <- df %>%
  dplyr::select(all_of(var.tot))%>%
  mutate(across(all_of(var.cont), as.numeric), 
         across(all_of(var.cat), as.factor)
         )

## Table 1 standard

table1 <- CreateTableOne(vars = var.tot, factorVars = var.cat, data =T1) 
table1 <- print(table1,
                nonnormal =var.cont,
                 contDigits = 1, showAllLevels = F)

table1_df <- as.data.frame(table1)
table1_df$Variables <- row.names(table1_df)

row.names(table1_df)<-NULL
table1_df <- table1_df[, c(2,1)]

T1_n <- T1%>%
  summarise(across(everything(), ~ sum(is.na(.))))%>%
  pivot_longer(cols= everything(),
    values_to = "n_missing",
    names_to = "ID")

table1_df1 <- table1_df %>%
  mutate(ID = sub("\\(median \\[IQR\\]\\)|= 1 \\(\\%\\)|= Women \\(\\%\\)|\\(\\%\\)|= Non ischémique \\(\\%\\)|= ischémique \\(\\%\\)", "",Variables),
         ID = trimws(ID))%>%
  mutate(ID =sub(" = YES","",ID))

table_1_vf <- left_join(table1_df1,T1_n, by="ID")%>%
  select(Variables,n_missing,Overall)

table_1_vf <- table_1_vf%>%
mutate(Variables= str_remove(Variables, " \\(median \\[IQR\\]\\)"))

traduction <- c(
  "n" = "n",
  "Gender = Women (%)" = "Gender = Female (%)",
  "HTN = 1 (%)" = "Hypertension (%)",
  "DM = 1 (%)" = "Diabetes Mellitus (%)",
  "Immunodepression = 1 (%)" = "Immunosuppression (%)",
  "CKD = 1 (%)" = "Chronic Kidney Disease (%)",
  "Neurologic_deficit = 1 (%)" = "Neurological Deficit (%)",
  "Chronic_respiratory_disease = 1 (%)" = "Chronic Respiratory Disease (%)",
  "HF = 1 (%)" = "Heart Failure (%)",
  "Cardiac_arrest_before_canul = 1 (%)" = "Cardiac Arrest Before Cannulation (%)",
  "cause = Non ischémique (%)" = "Non ichemic (%)",
  "ische = ischémique (%)" = "Ischemic (%)",
  "IABP_D0 = 1 (%)" = "Intra-Aortic Balloon Pump (%)",
  "Intubation = 1 (%)" = "Intubation (%)",
  "EER_D0 = 1 (%)" = "Renal Replacement Therapy (%)",
  "Outcome (%)" = "Outcome (%)",
  "   Bridge to LVAD" = "Bridge to LVAD",
  "   Bridge to Transplant" = "Bridge to Transplant",
  "   Death" = "Death in ICU",
  "   ECMO Weaning" = "ECMO Weaning",
  "Betablocker = YES (%)" = "Beta-blocker (%)",
  "Alphablocker = 1 (%)" = "Alpha-blocker (%)",
  "Levosimendan = YES (%)" = "Levosimendan (%)",
  "j90_deces = 1 (%)" = "90-day Mortality (%)",
  "j28_deces = 1 (%)" = "28-day Mortality (%)",
  "Age" = "Age (years)",
  "BMI" = "Body Mass Index (kg/m²)",
  "PAS_D0" = "Systolic Blood Pressure (mmHg)",
  "PAD_D0" = "Diastolic Blood Pressure (mmHg)",
  "PAM_D0" = "Mean Arterial Pressure (mmHg)",
  "Pulsepressure_D0" = "Pulse Pressure (mmHg)",
  "HR_D0" = "Heart Rate (bpm)",
  "Temp_D0" = "Temperature (°C)",
  "Preecmo_tte_ef" = "Pre-ECMO Ejection Fraction (%)",
  "Preecmo_tte_vtiao" = "Pre-ECMO Aortic Velocity Time Integral (cm)",
  "mean_NAD_24H" = "mean Norepinephrine Dose on cannulation day (gamma/kg/min)",
  "mean_DOB_24H" = "mean Dobutamine Dose on cannulation day (gamma/kg/min)",
  "VIS_D0" = "Vasoactive-Inotropic Score (VIS) at Day 0",
  "pH_D0" = "pH",
  "Sao2_D0" = "Oxygen Saturation (%)",
  "Paco2_D0" = "PaCO2 (mmHg)",
  "Lact_D0" = "Lactate (mmol/L)",
  "Hco3_D0" = "Bicarbonate (mmol/L)",
  "Creat_D0" = "Creatinine (µmol/L)",
  "Urea_D0" = "Urea (mg/dL)",
  "ALAT_D0" = "ALT (IU/L)",
  "ASAT_D0" = "AST (IU/L)",
  "Bili_D0" = "Bilirubin (mg/dL)",
  "PT_D0" = "Prothrombin Time (s)",
  "Tropo_i_hs_D0" = "High Sensitivity Troponin I (ng/L)",
  "Ntprobnp_D0" = "NT-proBNP (pg/mL)",
  "DPP3_D0" = "DPP3 (ng/mL)",
  "Plq_D0" = "Platelet Count (10⁹/L)",
  "Lenght_eer" = "Duration of Renal Replacement Therapy (days)",
  "Lenght_vm" = "Duration of Mechanical Ventilation (days)",
  "Time_hosp" = "ICU Length of Stay (days)",
  "Time_ecmo" = "ECMO Duration (days)"
)

table_1_vf <- table_1_vf %>%
  mutate(Variables = recode(Variables, !!!traduction))

order <- c(
  "n",
  # Demographics
  "Age (years)",
  "Gender = Female (%)",
  "Body Mass Index (kg/m²)",

  # Comorbidities
  "Hypertension (%)",
  "Diabetes Mellitus (%)",
  "Heart Failure (%)",
  "Chronic Kidney Disease (%)",
  "Chronic Respiratory Disease (%)",
  "Neurological Deficit (%)",
  "Immunosuppression (%)",
  
  #Cause
  "Non ichemic (%)",
  "Ischemic (%)",

  # Clinical presentation
  "Systolic Blood Pressure (mmHg)",
  "Diastolic Blood Pressure (mmHg)",
  "Mean Arterial Pressure (mmHg)",
  "Pulse Pressure (mmHg)",
  "Heart Rate (bpm)",
  "Temperature (°C)",
  "Cardiac Arrest Before Cannulation (%)",
  "Intra-Aortic Balloon Pump (%)",
  "Intubation (%)",
  "Renal Replacement Therapy (%)",

  # Echocardiography
  "Pre-ECMO Ejection Fraction (%)",
  "Pre-ECMO Aortic Velocity Time Integral (cm)",

  # Pharmacologic support
  "Beta-blocker (%)",
  "Alpha-blocker (%)",
  "mean Norepinephrine Dose on cannulation day (gamma/kg/min)",
  "mean Dobutamine Dose on cannulation day (gamma/kg/min)",
  "Vasoactive-Inotropic Score (VIS) at Day 0",
  "Levosimendan (%)",

  # Laboratory
  "pH",
  "Oxygen Saturation (%)",
  "PaCO2 (mmHg)",
  "Lactate (mmol/L)",
  "Bicarbonate (mmol/L)",
  "Creatinine (µmol/L)",
  "Urea (mg/dL)",
  "ALT (IU/L)",
  "AST (IU/L)",
  "Bilirubin (mg/dL)",
  "Prothrombin Time (s)",
  "High Sensitivity Troponin I (ng/L)",
  "NT-proBNP (pg/mL)",
  "DPP3 (ng/mL)",
  "Platelet Count (10⁹/L)",

  # Outcomes
  "Outcome (%)",
  "Death in ICU",
  "Bridge to LVAD",
  "Bridge to Transplant",
  "ECMO Weaning",
  "Duration of Renal Replacement Therapy (days)",
  "Duration of Mechanical Ventilation (days)",
  "ECMO Duration (days)",
  "ICU Length of Stay (days)",
  "28-day Mortality (%)",
  "90-day Mortality (%)"
)

table_1_vf2 <- table_1_vf %>%
  mutate(Variables = factor(Variables, levels = order)) %>%
  arrange(Variables)%>%
  rename(Missing = n_missing)

table1 <- flextable(as.data.frame(table_1_vf2)) %>% autofit() %>%
      set_caption("Table 1: Characteristics of the population") %>% 
     align(j = c(2:3), align = "center", part = "all")

```

```{r, include=FALSE, eval=F}
doc <- read_docx() %>%
  body_add_flextable(table1)

print(doc, target = "TABLES/Table1_def.docx")
```


```{r}
table1
```

# Table 2a: Hemodynamic parameters on weaning day

```{r, include=FALSE}
## General information at weaning day
df_long <- df %>%
  select(
    ID,
    Time_weaning,
    jsev_ecmo_debit,
    jsev_dobu_cum,
    jsev_NAd_cum,
    jsev_pam,
    jsev_freq,
    jsev_ett_fevg,
    jsev_ett_vmax_s_mit_lat_vg,
    jsev_lact
  ) %>%
  rename(
    `delay to weaning test` = Time_weaning,
    `ECMO flow (l)` = jsev_ecmo_debit,
    `Dobutamine cumulative (mg)` = jsev_dobu_cum,
    `Noradrenaline cumulative (mg)` = jsev_NAd_cum,
    `MAP_baseline (mmHg)` = jsev_pam,
    `HR (bpm)` = jsev_freq,
    `FEVG (%)` = jsev_ett_fevg,
    `Mitral S' (cm/s)` = jsev_ett_vmax_s_mit_lat_vg,
    `Lactate (mmol/L)` = jsev_lact
  ) %>%
  pivot_longer(
    cols = -ID,
    names_to = "Parameter",
    values_to = "Value"
  )

# Define desired parameter order
custom_order <- c(
  "delay to weaning test",
  "ECMO flow (l)",
  "Dobutamine cumulative (mg)",
  "Noradrenaline cumulative (mg)",
  "MAP_baseline (mmHg)",
  "HR (bpm)",
  "FEVG (%)",
  "Mitral S' (cm/s)",
  "Lactate (mmol/L)"
)

# Apply ordering after summarising
table2a <- df_long %>%
  group_by(Parameter) %>%
  summarise(
    n = sum(!is.na(Value)),
    Median = round(median(Value, na.rm = TRUE), 1),
    Q1 = round(quantile(Value, 0.25, na.rm = TRUE), 1),
    Q3 = round(quantile(Value, 0.75, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  mutate(`Median [IQR]` = paste0(Median, " [", Q1, "–", Q3, "], n=", n)) %>%
  select(Parameter, `Median [IQR]`) %>%
  slice(match(custom_order, Parameter))  # Reorder according to custom_order
```

```{r}
table2a
```

# Table 2b  hemodynamic parameters during the weaning test

```{r,INCLUDE=F}
# Create new variables for flow condition and unify variable names

df_long <- df %>%
  select(
    ID,
    jsev_epr_sev_t0_3l_itv,
    jsev_epr_sev_t0_3l_pam,
    `jsev_epr_sev_t1_1,5l_itv`,
    `jsev_epr_sev_t1_1,5l_pam`
  ) %>%
  rename(
    `VTI (cm)` = jsev_epr_sev_t0_3l_itv,
    `MAP (mmHg)` = jsev_epr_sev_t0_3l_pam,
    `VTI_1.5L (cm)` = `jsev_epr_sev_t1_1,5l_itv`,
    `MAP_1.5L (mmHg)` = `jsev_epr_sev_t1_1,5l_pam`,
  ) %>%
  pivot_longer(
    cols = -c(ID),
    names_to = "Parameter_raw",
    values_to = "Value"
  ) %>%
  mutate(
    Value = as.numeric(Value),
    Condition = case_when(
      str_detect(Parameter_raw, "1.5L") ~ "1.5L",
      TRUE ~ "3L"
    ),
    Condition = factor(Condition, levels = c("3L", "1.5L")),
    Parameter = str_replace(Parameter_raw, "_1.5L", "")
  )

# Create summary table

table2b <- df_long %>%
  group_by(Parameter, Condition) %>%
  summarise(
    n = sum(!is.na(Value)),
    Median = round(median(Value, na.rm = TRUE), 5),
    Q1 = round(quantile(Value, 0.25, na.rm = TRUE), 5),
    Q3 = round(quantile(Value, 0.75, na.rm = TRUE), 5),
    .groups = "drop"
  ) %>%
  mutate(
    Value_IQR = paste0(Median, " [", Q1, "–", Q3, "]; n=", n)
  ) %>%
  select(Parameter, Condition, Value_IQR) %>%
  pivot_wider(names_from = Condition, values_from = Value_IQR)

TEST<- wilcox.test(Value ~ Condition,
            data = df_long %>% filter(Parameter == "VTI (cm)"))
```


```{r}
table2b
```

No significant difference between VTI at 3 L and 1.5 L flow conditions at weaning day (Wilcoxon test, p = `r format.pval(TEST$p.value, digits=3)`)

# Figure 1: Association between ADRB expressions and outcomes

```{r,include=FALSE}

prep_A1 <- prep_marker_data(df, prefix = "ADR1L", log1p_y = F)
test_A1 <- test_marker(prep_A1$data)
Figure_1_1 <- plot_marker(
  dat      = prep_A1$data,
  counts   = prep_A1$counts,
  x_labels = prep_A1$x_labels,
  panel_title = "A",
  y_var    = "value_log",
  y_lab    = "Percentage of T4 lymphocyte expressing ADRB1\n(log1p scale)"
) +
  annotate("text",
           x = 1,
           y = min(prep_A1$data$value_log, na.rm = TRUE) +1,
           label = test_A1$label,
           hjust = 0)


# ADRB2 (lymphocytes) — log1p pour modèle & plot
prep_A2 <- prep_marker_data(df, prefix = "ADR2L", log1p_y = F)
test_A2 <- test_marker(prep_A2$data)
Figure_1_2 <- plot_marker(
  dat      = prep_A2$data,
  counts   = prep_A2$counts,
  x_labels = prep_A2$x_labels,
  panel_title = "B",
  y_var    = "value_log",
  y_lab    = "Percentage of T4 lymphocyte expressing ADRB2\n(log1p scale)"
) +
  annotate("text",
           x = 1,
           y = min(prep_A2$data$value_log, na.rm = TRUE) +1,
           label = test_A2$label,
           hjust = 0)

# ADRB1 (monocytes) — on peut tracer en brut mais le modèle reste sur log1p
prep_M1 <- prep_marker_data(df, prefix = "M_ADR1", log1p_y = F)
test_M1 <- test_marker(prep_M1$data)
Figure_1_3 <- plot_marker(
  dat      = prep_M1$data,
  counts   = prep_M1$counts,
  x_labels = prep_M1$x_labels,
  panel_title = "C",
  y_var    = "value_log",   # tracé en valeurs brutes si tu préfères
  y_lab    = "Percentage of monocytes expressing ADRB1\n(log1p scale)"
) +
  annotate("text",
           x = 2,
           y = min(prep_M1$data$value, na.rm = TRUE),
           label = test_M1$label,
           hjust = 0)

# ADRB2 (monocytes)
prep_M2 <- prep_marker_data(df, prefix = "M_ADR2", log1p_y = F)
test_M2 <- test_marker(prep_M2$data)
Figure_1_4 <- plot_marker(
  dat      = prep_M2$data,
  counts   = prep_M2$counts,
  x_labels = prep_M2$x_labels,
  panel_title = "D",
  y_var    = "value_log",
  y_lab    = "Percentage of monocytes expressing ADRB2\n(log1p scale)",
   show_legend = TRUE   # 👈 ici seule cette figure garde la légende
) +
  annotate("text",
           x = 1,
           y = min(prep_M2$data$value, na.rm = TRUE)+1,
           label = test_M2$label,
           hjust = 0)

F1_4<- prep_M2$data %>% filter(!is.finite(value_log)) %>% count(time, Outcomes)


# Combine figures

Figure1_tot <- (Figure_1_1 + Figure_1_2 + Figure_1_3 + Figure_1_4) +
  patchwork::plot_layout(guides = "collect") &
  theme(legend.position = "bottom")   # ou "right"
```

```{r Figure1_tot, fig.width=8, fig.height=6, fig.scale=1.5, dpi=300, fig.align='center'}
Figure1_tot
```

# Figure 2: Association between VTI and ADRB in Lymphocyte at admission

```{r,include=TRUE}
#ADRB1

df$ADR1L_J0
F2_1_a <- df %>%
select(`jsev_epr_sev_t1_1,5l_itv`, ADR1L_J0)%>%
  mutate(
    VTI_JS = case_when(
      `jsev_epr_sev_t1_1,5l_itv` > 12 ~ ">12 cm",
      `jsev_epr_sev_t1_1,5l_itv` <= 12 ~ "≤12 cm"
    )
  )%>%
  na.omit()
   

 # Mann-Whitney test for VTI and ADRB1
mann_whitney_VTI <- wilcox.test(log1p(ADR1L_J0) ~ VTI_JS, data = F2_1_a)

# Extract and format the p-value
p_value_VTI <- sprintf("%.3f", mann_whitney_VTI$p.value)

count_data <- F2_1_a %>%
  group_by(VTI_JS) %>%
  summarise(n = n())

# Boxplot for ADRB1 with p-value annotation 
Figure2a <- ggplot(F2_1_a, aes(x = VTI_JS, y = log1p(ADR1L_J0), fill = VTI_JS)) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7, position = position_jitter(width = 0.1)) +
  geom_text(data = count_data,
            aes(x = VTI_JS, y = min(log1p(F2_1_a$ADR1L_J0), na.rm = TRUE) * 1.25,
                label = paste0("n=", n)),
            size = 4,
            inherit.aes = FALSE) +
  scale_fill_manual(values = c("≤12 cm" = "#0072B2", ">12 cm" = "#D55E00")) +
  labs(
    title = "A",
    x = "VTI measured at 1.5 L/min",
    y = "% of lymphocytes expressing ADRB1 at ECMO weaning \n (log scale)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  annotate("text",
           x = 1.5,
           y = max(log1p(F2_1_a$ADR1L_J0), na.rm = TRUE),
           label = paste("p:", p_value_VTI),
           size = 5, color = "black", hjust = 0.5)


#ADRB2
F2_2_a <- df %>%
select(`jsev_epr_sev_t1_1,5l_itv`, ADR2L_J0)%>%
  mutate(
    VTI_JS = case_when(
      `jsev_epr_sev_t1_1,5l_itv` > 12 ~ ">12 cm",
      `jsev_epr_sev_t1_1,5l_itv` <= 12 ~ "≤12 cm"
    )
  )%>%
  na.omit()
   

 # Mann-Whitney test for VTI and ADRB2
mann_whitney_VTI_2 <- wilcox.test(log1p(ADR2L_J0) ~ VTI_JS, data = F2_2_a)

# Extract and format the p-value
p_value_VTI_2 <- sprintf("%.3f", mann_whitney_VTI_2$p.value)

count_data <- F2_2_a %>%
  group_by(VTI_JS) %>%
  summarise(n = n(), .groups = "drop")

# Boxplot for ADRB2 with p-value annotation 
Figure2b <- ggplot(F2_2_a, aes(x = VTI_JS, y = log1p(ADR2L_J0), fill = VTI_JS)) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7, position = position_jitter(width = 0.1)) +
  scale_fill_manual(values = c("≤12 cm" = "#0072B2", ">12 cm" = "#D55E00")) +
  labs(
    title = "B",
    x = "VTI measured at 1.5L /min",
    y = "% of lymphocytes expressing ADRB2 measured at ECMO weaning \n (log10 scale)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  annotate(
    "text",
    x = 1.5,
    y = max(log1p(F2_2_a$ADR2L_J0), na.rm = TRUE),
    label = paste("p:", p_value_VTI_2),
    size = 5,
    color = "black",
    hjust = 0.5
  ) +
  geom_text(
    data = count_data,
    aes(
      x = VTI_JS,
      y = min(log1p(F2_2_a$ADR2L_J0), na.rm = TRUE) * 1.25,  
      label = paste0("n=", n)
    ),
    inherit.aes = FALSE,
    size = 4
  )

Figure2_tot <- Figure2a | Figure2b +
  plot_annotation(
    title = "Association between VTI at weaning and ADRB1/ADRB2 expression in lymphocyte",
    tag_levels = c("A","B")  
  )

```

```{r}
Figure2_tot
```

# Figure 3: Association between VTI and % of expression of ADRB on monocyte at admission

```{r,include=TRUE}
#ADRB1

F3_1 <- df %>%
select(`jsev_epr_sev_t1_1,5l_itv`, M_ADR1_J0)%>%
  mutate(
    VTI_JS = case_when(
      `jsev_epr_sev_t1_1,5l_itv` > 12 ~ ">12 cm",
      `jsev_epr_sev_t1_1,5l_itv` <= 12 ~ "≤12 cm"
    )
  )%>%
  na.omit()
   
 # Mann-Whitney test for VTI and ADRB1
mann_whitney_VTI <- wilcox.test(log1p(M_ADR1_J0) ~ VTI_JS, data = F3_1)

# Extract and format the p-value
p_value_VTI <- sprintf("%.3f", mann_whitney_VTI$p.value)

count_data <- F3_1 %>%
  group_by(VTI_JS) %>%
  summarise(n = n())

# Boxplot for ADRB1 with p-value annotation 
Figure3a <- ggplot(F3_1, aes(x = VTI_JS, y = log1p(M_ADR1_J0), fill = VTI_JS)) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7, position = position_jitter(width = 0.1)) +
  geom_text(data = count_data,
            aes(x = VTI_JS, y = -3,
                label = paste0("n=", n)),
            size = 4,
            inherit.aes = FALSE) +
  scale_fill_manual(values = c("≤12 cm" = "#0072B2", ">12 cm" = "#D55E00")) +
  labs(
    title = "A",
    x = "VTI measured at 1.5 L/min",
    y = "% of Monocytes expressing ADRB1  measured at ECMO weaning \n (log1 scale)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  annotate("text",
           x = 1.5,
           y = max(log1p(F3_1$M_ADR1_J0), na.rm = TRUE),
           label = paste("p:", p_value_VTI),
           size = 5, color = "black", hjust = 0.5)


#ADRB2
F3_2 <- df %>%
select(`jsev_epr_sev_t1_1,5l_itv`, M_ADR2_J0)%>%
  mutate(
    VTI_JS = case_when(
      `jsev_epr_sev_t1_1,5l_itv` > 12 ~ ">12 cm",
      `jsev_epr_sev_t1_1,5l_itv` <= 12 ~ "≤12 cm"
    )
  )%>%
  na.omit()
   

 # Mann-Whitney test for VTI and ADRB2
mann_whitney_VTI_2 <- wilcox.test(log1p(M_ADR2_J0) ~ VTI_JS, data = F3_2)

# Extract and format the p-value
p_value_VTI_2 <- sprintf("%.3f", mann_whitney_VTI_2$p.value)

count_data <- F3_2 %>%
  group_by(VTI_JS) %>%
  summarise(n = n(), .groups = "drop")

# Boxplot for ADRB2 with p-value annotation 
Figure3b <- ggplot(F3_2, aes(x = VTI_JS, y = log1p(M_ADR2_J0), fill = VTI_JS)) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7, position = position_jitter(width = 0.1)) +
  scale_fill_manual(values = c("≤12 cm" = "#0072B2", ">12 cm" = "#D55E00")) +
  labs(
    title = "B",
    x = "VTI measured at 1.5L /min",
    y = "% of Monocytes expressing ADRB2 at ECMO weaning \n (log scale)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  annotate(
    "text",
    x = 1.5,
    y = max(log1p(F3_2$M_ADR2_J0), na.rm = TRUE),
    label = paste("p:", p_value_VTI_2),
    size = 5,
    color = "black",
    hjust = 0.5
  ) +
  geom_text(
    data = count_data,
    aes(
      x = VTI_JS,
      y = min(log1p(F3_2$M_ADR2_J0), na.rm = TRUE) * 1.25,  
      label = paste0("n=", n)
    ),
    inherit.aes = FALSE,
    size = 4
  )

Figure3_tot <- Figure3a | Figure3b +
  plot_annotation(
    title = "Association between VTI at weaning and % of monocytes expressing  ADRB",
    tag_levels = c("A","B")  
  )

```

```{r}
Figure3_tot
```
#Figure4 :Association between IL6 at admission and VTI

```{r}
F4_1 <- df %>%
select(`jsev_epr_sev_t1_1,5l_itv`, IL6_J0)%>%
  mutate(
    VTI_JS = case_when(
      `jsev_epr_sev_t1_1,5l_itv` > 12 ~ ">12 cm",
      `jsev_epr_sev_t1_1,5l_itv` <= 12 ~ "≤12 cm"
    )
  )%>%
  na.omit()
   

 # Mann-Whitney test for VTI and ADRB2
mann_whitney_VTI_3 <- wilcox.test(log1p(IL6_J0) ~ VTI_JS, data = F4_1)

# Extract and format the p-value
p_value_VTI_2 <- sprintf("%.3f", mann_whitney_VTI_2$p.value)

count_data <- F4_1 %>%
  group_by(VTI_JS) %>%
  summarise(n = n(), .groups = "drop")
# Boxplot for ADRB2 with p-value annotation 
Figure4 <- ggplot(F3_3, aes(x = VTI_JS, y = log1p(IL6_J0), fill = VTI_JS)) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7, position = position_jitter(width = 0.1)) +
  scale_fill_manual(values = c("≤12 cm" = "#0072B2", ">12 cm" = "#D55E00")) +
  labs(
    title = "B",
    x = "VTI measured at 1.5L /min",
    y = "IL6  at implantation \n (log scale)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  annotate(
    "text",
    x = 1.5,
    y = max(log1p(F4_1$IL6_J0), na.rm = TRUE),
    label = paste("p:", p_value_VTI_2),
    size = 5,
    color = "black",
    hjust = 0.5
  ) +
  geom_text(
    data = count_data,
    aes(
      x = VTI_JS,
      y = min(log1p(F4_1$IL6_J0), na.rm = TRUE) * 1.25,  
      label = paste0("n=", n)
    ),
    inherit.aes = FALSE,
    size = 4
  )

```


# Figure 5: Cytokines

# IL-6

```{r}
plot_cytokine(df, "IL6")
```

# IL-5

```{r}
plot_cytokine(df, "IL5")
```


# IL-8

```{r}
plot_cytokine(df, "IL8/CXCL8")
```


# IL-10

```{r}
plot_cytokine(df, "IL10")
```

# GDF-15

```{r}
plot_cytokine(df, "GDF-15")
```

```{r}

F5_1 <- df %>%
select(`jsev_epr_sev_t1_1,5l_itv`, L_TH2_J0)%>%
  mutate(
    VTI_JS = case_when(
      `jsev_epr_sev_t1_1,5l_itv` > 12 ~ ">12 cm",
      `jsev_epr_sev_t1_1,5l_itv` <= 12 ~ "≤12 cm"
    )
  )%>%
  na.omit()
   

 # Mann-Whitney test for VTI and ADRB2
mann_whitney_VTI_5 <- wilcox.test(log1p(L_TH2_J0) ~ VTI_JS, data = F5_1)

# Extract and format the p-value
p_value_VTI_5 <- sprintf("%.3f", mann_whitney_VTI_5$p.value)

count_data <- F5_1 %>%
  group_by(VTI_JS) %>%
  summarise(n = n(), .groups = "drop")
# Boxplot for TH2 with p-value annotation 

Figure3d <- ggplot(F5_1, aes(x = VTI_JS, y = log1p(L_TH2_J0), fill = VTI_JS)) +
  geom_boxplot() +
  geom_point(size = 2, alpha = 0.7, position = position_jitter(width = 0.1)) +
  scale_fill_manual(values = c("≤12 cm" = "#0072B2", ">12 cm" = "#D55E00")) +
  labs(
    title = "B",
    x = "VTI measured at 1.5L /min",
    y = "% of LT TH2 at implantation \n (log scale)"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  annotate(
    "text",
    x = 1.5,
    y = max(log1p(F5_1$L_TH2_J0), na.rm = TRUE),
    label = paste("p:", p_value_VTI_5),
    size = 5,
    color = "black",
    hjust = 0.5
  ) +
  geom_text(
    data = count_data,
    aes(
      x = VTI_JS,
      y = min(log1p(F5_1$L_TH2_J0), na.rm = TRUE) * 1.25,  
      label = paste0("n=", n)
    ),
    inherit.aes = FALSE,
    size = 4
  )
```


# Figure 6: Survival \~ IL6 admission

```{r,include=FALSE}
df <- df%>%
  mutate(IL6_median = as.factor(if_else(df$IL6_J0<=median(df$IL6_J0), "<=median", ">median" )
                                ),
         outcome_censored = case_when(Outcome=="Bridge to LVAD"|Outcome=="Bridge to Transplant"~0,
                                Outcome=="ECMO Weaning"~ 0, T~1),
         outcome_FG = factor(case_when(Outcome=="Bridge to LVAD"|Outcome=="Bridge to Transplant"~2,
                                Outcome=="ECMO Weaning"~ 0, T~1))
  )  

#IL6
km_fit <- survfit(Surv(diff_days, j90_deces) ~ IL6_median , data = df)
summary(km_fit)

Figure_6 <- ggsurvplot(km_fit, data = df, 
  pval = TRUE,               
  pval.coord = c(10, 0.1),   
  conf.int = FALSE,           
  risk.table = TRUE,         
  risk.table.height = 0.2,   
  risk.table.y.text = TRUE,  
  risk.table.title = "Number at Risk",  
  xlim = c(0, 90),
  ylab = "90 day survival",
  break.time.by = 7,
  legend.title = " ",       
  palette = c("#E63946", "#457B9D"),
  ggtheme = theme_minimal(),
  legend.labs = c("below median", "above median"),
  legend.position= "none"
)

####MODELE CENSURE POUR TRANSPLANTATION ET LVAD 

fg <- finegray(Surv(diff_days, outcome_FG) ~ IL6_median, data = df, etype = 1)
fit_fg <- coxph(Surv(fgstart, fgstop, fgstatus) ~ IL6_median,
                data = fg, weights = fgwt)
summary(fit_fg)

km_fit_comp <- survfit(Surv(diff_days, outcome_censored) ~ IL6_median , data = df)
summary(km_fit)

mod_censored <-coxph(Surv(diff_days, outcome_censored) ~ IL6_median , data = df)

summary(mod_censored)# similar

library(broom)

res_cox <- tidy(mod_censored, exponentiate = TRUE, conf.int = TRUE)
lab <- res_cox %>%
  filter(grepl("^IL6_median", term)) %>%
  mutate(
    label = sprintf(
      "HR = %.2f (95%% CI %.2f–%.2f)",
      estimate, conf.low, conf.high
    )
  ) %>%
  pull(label)

lab

Figure_6_censored <- ggsurvplot(km_fit_comp, data = df, 
  pval = F,               
  conf.int = FALSE,           
  risk.table = TRUE,         
  risk.table.height = 0.2,   
  risk.table.y.text = TRUE,  
  risk.table.title = "Number at Risk",  
  xlim = c(0, 90),
  ylab = "90 day survival",
  break.time.by = 7,
  legend.title = " ",       
  palette = c("#E63946", "#457B9D"),
  ggtheme = theme_bw(),
  legend.labs = c("below median", "above median"),
  legend.position= "none"
)

xmax <- max(df$diff_days, na.rm = TRUE)
Figure_6_censored$plot <- Figure_6_censored$plot +
  annotate("text", x = 0.2 * xmax, y = 0.95, hjust = 0,
           label = lab, size = 3.7)

```

```{r}
Figure_6_censored
```

#Figure 7: Association between IL6 and vasopressor dose

```{r}

F7 <- df%>%
  select(IL6_J0, NADcum_D0)
        
# Nonparametric correlation (Spearman)
cor_test_result <- cor.test(
  df$IL6_J0, df$NADcum_D0,
  method = "spearman",
  use = "complete.obs"
)

# Print correlation test
cor_test_result

# Extract rho and p-value for annotation
rho  <- round(unname(cor_test_result$estimate), 2)
pval <- signif(cor_test_result$p.value, 2)

df <- df %>%
  mutate(rank_IL6 = rank(IL6_J0, ties.method = "average"),
         rank_NAD = rank(NADcum_D0, ties.method = "average"))

ggplot(df, aes(x = rank_IL6, y = rank_NAD)) +
  geom_point(alpha = 0.7, color = "steelblue") +
  geom_smooth(method = "loess", se = TRUE, color = "darkred") +
  labs(
    title = "Rank–rank association (Spearman)",
    x = "IL-6 rank",
    y = "Cumulative norepinephrine dose rank"
  ) +
  theme_classic(base_size = 14)+
  annotate(
    "text", x = Inf, y = Inf,
    label = paste0("Spearman ρ = ", rho, "\nP = ", pval),
    hjust = 1.1, vjust = 1.5, size = 4.5
  )
```


